---
title: "1-Data processing"
author: "Aurore Ponchon"
format: html
editor: visual
---

```{r}
#| echo: false
#| include: true
#| warning: false
#| message: false


library(move)
library(here)
library(sf)
library(tidyverse)
library(lubridate)
library(conflicted)
library(readxl)
library(grid)
library(gridExtra)
library(sf)
library(rnaturalearth)
library(kableExtra)

source(here::here("R","trip_functions.R"))
```

# 1-Importing tracking data

Data are deposited on movebank.com. An account is required to access and download the data.

```{r}
#| echo: false
#| include: true
#| warning: false
#| message: false


#Import data directly from movebank
# loginStored <- movebankLogin(username="AurorePonchon",
#                                password="***")
# 
# gannets<-searchMovebankStudies(x="gannet",
#                                  login=loginStored)[1]

# metadata<-getMovebankStudy(study=gannets,
#                            login=loginStored)
# getMovebankSensorsAttributes(study=gannets,
#                              login=loginStored)
# reftable<-getMovebankReferenceTable(study=gannets,
#                                     login=loginStored)

# #raw data from 2022 and 2023 are downloaded and saved
# move.rouzic <- getMovebankData(study=gannets,
#                                 login=loginStored,
#                               removeDuplicatedTimestamps=TRUE)
# save(move.rouzic,
#       file=here("data","RawData","gannets_rouzic_raw_2022_2023.Rdata"))


#raw data from 2022 nd 2023 
load(here("data","RawData","gannets_rouzic_raw_2022_2023.Rdata"))

 #format raw data and select only useful columns
all_rouzic<-move.rouzic@data %>% 
    dplyr::select(tag_local_identifier,
                  timestamp,
                  location_long,
                  location_lat,
                  height_above_msl, 
                  ground_speed) %>% 
    rename(id=tag_local_identifier,
           datetime=timestamp,
           long=location_long,
           lat=location_lat,
           alt=height_above_msl,
           speed=ground_speed)  %>% 
    arrange(id,datetime) %>% 
    mutate(id=as.factor(id),
           site=as.factor("Rouzic"),
           year=year(datetime))

```

# 2-Importing individual data

Individual data (sex, date of equipment, breeding stage, period, serological status, iris colour etc...) are stored in a separate file. Both jueveniles and breeding aduts are part of the dataset.

```{r}
#| echo: false
#| include: true
#| warning: false
#| message: false

serol.rouzic<-get_serol_data_2023()
```

# 3- Classifying individual at-sea trips

Locations are classified in consecutive trips at-sea when the birds leave the colony further than 1km away from the colony and for more than 1h. Preliminary plots are used to identify the duration of time gaps.

```{r}
#| echo: false
#| include: true
#| warning: false
#| message: false

all_rouzic23<-all_rouzic %>% 
              dplyr::filter(year==2023) %>% 
              droplevels() %>% 
              left_join(., serol.rouzic, by="id",keep=F)
  
remove(all_rouzic)  

ids<-unique(levels(all_rouzic23$id))
colo.ind<-tim.colors(length(ids))

#create new object with distance and time differences
raw_rouzic_2023<-NULL

#loop through each individual 
for (i in 1:length(ids)){
    ind.loc<-all_rouzic23 %>% 
        dplyr::filter(id==ids[i]) %>% 
        mutate(distmax=0,
               distadj=0,
               totalpath=0,
               difftimemin=0,
               travelNb=0,
               onlandNb=0)
  
    print(c("Processing individual ", i))
      
  #loop through each line to calculate distance and time difference with previous location
for (l in 2:nrow(ind.loc)){
    
ind.loc[l,"distadj"]<-rdist.earth(ind.loc[l,c("long","lat")],
                                   ind.loc[l-1,c("long","lat")], 
                                  miles = F)
ind.loc[l,"difftimemin"]<-difftime(ind.loc[l,"datetime"],
                                   ind.loc[l-1,"datetime"], 
                                   units="mins")
ind.loc[l,"distmax"]<-rdist.earth(ind.loc[l,c("long","lat")],
                                   colo_coord_rouzic[c(1,2)], 
                                  miles = F)
}

#define periods on land and at sea   
ind.loc<-define_trips(data=ind.loc,
                      dist.min=dist.threshold)
raw_rouzic_2023<-rbind(raw_rouzic_2023,ind.loc)    
} #end of loop on individuals


raw_rouzic_2023$trip.id<-paste(raw_rouzic_2023$id,raw_rouzic_2023$travelNb,sep=".")
raw_rouzic_2023<-add_missing_return(raw_rouzic_2023,
                                  time.int=reso,colony=colo_coord_rouzic)

save(raw_rouzic_2023,file=here("data","NewlyCreatedData","rawLocRouzic2023.Rdata"))

load(here("data","NewlyCreatedData","rawLocRouzic2023.Rdata"))

#plot distance to the colony over time
distraw<-ggplot(raw_rouzic_2023,aes(x=datetime,y=distmax),group=id)+
    geom_point(aes(color=id),show.legend=F) +
    scale_y_continuous(limits=c(0,200)) +
    scale_color_manual(values=colo.ind)+
    labs(x="Date",y="Maximal distance to colony (km)")

#plot time gaps
diffraw<-ggplot(raw_rouzic_2023,aes(x=datetime,y=difftimemin),group=id)+
    geom_point(aes(color=id),show.legend=F) +
    scale_y_continuous(limits=c(0,1500)) +
    scale_color_manual(values=colo.ind)+
    labs(x="Date",y="Time gaps between 2 locations (in min)")

grid.arrange(distraw,diffraw, ncol=2,
              top = grid::textGrob("Raw data",gp=gpar(fontsize=16,font=3)))

#get overview of trips for each individual
# 
# target<-c("France","United Kingdom","Spain","Morocco","Belgium",
#           "Netherlands","Ireland")
# 
# #contours of countries
# shp<-ne_countries(scale = 10, returnclass = 'sf') %>%
#    dplyr::filter(name %in% target) 
# 
# for (a in 1:length(ids)){
#     
#     #extract data for 1 ind
#      indi<-raw_rouzic_2023 %>% 
#         dplyr::filter(id==ids[a])
#     
#      #get sf layer for 1 ind
#     indisf<-raw_rouzic_2023 %>% 
#         dplyr::filter(id==ids[a]) %>% 
#         st_as_sf(.,coords=c("long","lat")) %>% 
#         st_set_crs(st_crs(4326))
#     
# #assign shape according to life stage
#     shap<-ifelse(indi$stage[1]=="juvenile",17,16)
# 
#     #create external files
#     png(here::here("outputs","Trips by individuals",
#                    paste0(ids[a],"_raw_trips.png")),
#         width=5000,height=3000,res=400)
# 
#     #plot of maximal distance to the colony    
#    distrawind<-ggplot(indi,aes(x=datetime,y=distmax))+
#     geom_point(aes(color=datetime),shape=shap,show.legend=F) +
#     labs(x="Date",y="Maximal distance (km)")
#  
#    #plot of time gaps
#    diffrawind<-ggplot(indi,aes(x=datetime,y=difftimemin,
#                                group=stage))+
#     geom_point(aes(color=datetime),shape=shap,
#                show.legend=F) +
#     labs(x="Date",y="Time gap (in min)")
#    
#    #Map of all locations for 1 ind
#    map<-ggplot()+
#        geom_sf(data=indisf, aes(color=datetime),
#                shape=shap,
#                show.legend=F) +
#        geom_sf(data=shp,fill="grey90") +
#        
#        coord_sf(xlim=c(floor(st_bbox(indisf)[1]),
#                        ceiling(st_bbox(indisf)[3])),
#                 ylim=c(floor(st_bbox(indisf)[2]),
#                        ceiling(st_bbox(indisf)[4]))) +
#   theme(panel.background = element_rect(fill = "lightblue"))
# 
#    grid.arrange(distrawind,diffrawind, map,ncol=3,
#               top = grid::textGrob(paste0("Raw data ",ids[a]),
#                                    gp=gpar(fontsize=14,font=3)))
#   dev.off()
#   
# }

#create shapefiles
SPDF <-st_as_sf(raw_rouzic_2023,
         coords = c("long", "lat"), 
        crs = 4326)

 st_write(SPDF,
           dsn=here("data","NewlyCreatedData","shapefiles"),
          layer="raw_trips_all_rouzic23",
          driver="ESRI Shapefile",
          delete_layer =T)


```

# 4- Summary of raw at-sea trips

Calculation of at-sea trip characteristics and periods on land

```{r}
#| echo: false
#| include: true
#| warning: false
#| message: false

load(here("data","NewlyCreatedData","rawLocRouzic2023.Rdata"))
ids<-unique(levels(raw_rouzic_2023$id))
colo.ind<-tim.colors(length(ids))


#summarize all trips and periods on land by individuals
raw_trips_rouzic23<-trips_summary_ind(raw_rouzic_2023,
                                      colony=colo_coord_rouzic) %>% 
                    dplyr::filter(Distmaxkm>1) 

raw_land_rouzic23<-land_summary_ind(raw_rouzic_2023,colony=colo_coord_rouzic)

#summarize all trips at the population level
summy.trips<-raw_trips_rouzic23 %>% 
    group_by(id) %>% 
    summarize(Nbtrip=n(),
              Mindurh=min(TripDurh),
              MaxDist=max(Distmaxkm),
              maxDiffh=max(maxDiffTimeh),
              minLocs=min(nlocs),
              NbLocs=sum(nlocs))

#details individual trips
kbl(summy.trips,caption="Summary of raw trips") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

#summarize all periods on land at the population level
summy.land<-raw_land_rouzic23 %>% 
    group_by(id) %>% 
    summarize(Nbland=n(),
              Mindurh=min(LandDurh),
              maxDiffh=max(maxDiffTimeh),
              minLocs=min(nlocs),
              NbLocs=sum(nlocs))

#details individual trips
kbl(summy.trips,caption="Summary of raw trips") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))


##plot trip and land periods
dist<-ggplot(raw_trips_rouzic23,aes(x=DateEnd,y=Distmaxkm,group=id))+
    geom_point(aes(color=id, shape=stage),show.legend=F) +
 #   scale_shape_manual(values=c(16,17)) +
    scale_y_continuous(limits=c(0,500)) +
    scale_color_manual(values=colo.ind) +
    labs(x="Date",y="Maximal distance to colony (km)")

dur<-ggplot(raw_trips_rouzic23,aes(x=DateEnd,y=TripDurh ,group=id))+
    geom_point(aes(color=id,shape=stage),show.legend=F) +
  #  scale_y_continuous(limits=c(0,10)) +
    scale_color_manual(values=colo.ind)+
    labs(x="Date",y="Trip duration (h")

correl<-ggplot(raw_trips_rouzic23,aes(x=log(TripDurh),y=log(Distmaxkm)))+
    geom_point(aes(color=DateEnd, shape=stage), show.legend=F) +
    scale_colour_datetime(low=viridis(100)[1],high=viridis(100)[100]) +
    labs(x="log Trip duration (h)",y="log Maximal distance to colony (km)")

land<-ggplot(raw_land_rouzic23,aes(x=DateEnd,y=LandDurh ,group=id))+
    geom_point(aes(color=id,shape=stage), show.legend=F) +
  #  scale_y_continuous(limits=c(0,10)) +
    scale_color_manual(values=colo.ind)+
    labs(x="Date",y="Periods on land (h)")

grid.arrange(dist,dur,correl, land, ncol=2, nrow=2,
              top = textGrob("Raw trip features",gp=gpar(fontsize=16,font=3)))

```

# 5- Filtering complete trips

```{r}
#| echo: false
#| include: true
#| warning: false
#| message: false


max.gap<-10

#filter trips based on defined threshold
clean.trip.ad.nb23<-raw_trips_rouzic23 %>%
     dplyr::filter(TripDurh > dur.thres & 
                       nlocs >= row.thres & 
                       stage=="adult" & 
                       maxDiffTimeh < max.gap) 


#get incomplete trips
incomp<-raw_rouzic_2023 %>% 
    group_by(trip.id) %>% 
    summarize(Max=max(datetime), Dist=dplyr::last(distmax)) %>% 
    dplyr::filter(Dist>1) 

#clean trips
cleanLocs.rouzic23<-raw_rouzic_2023 %>%
    dplyr::filter(trip.id %in% 
                      clean.trip.ad.nb23$trip.id) %>% 
    dplyr::filter(!trip.id %in% incomp$trip.id)


save(cleanLocs.rouzic23,
     file=here::here("data","NewlyCreatedData",
               paste0("clean_",max.gap,"_locs_AD_rouzic23.Rdata")))

#save as shapefiles
SPDF <-st_as_sf(cleanLocs.rouzic23,
         coords = c("long", "lat"), 
        crs = 4326)

 st_write(SPDF,
           dsn=here("data","NewlyCreatedData","shapefiles"),
          layer=paste0("clean_",max.gap,"_locs_AD_rouzic23"),
          driver="ESRI Shapefile",
          delete_layer =T)
 

```

# 6- Interpolation

```{r}
#| echo: false
#| include: true
#| warning: false
#| message: false


load(here("data","NewlyCreatedData",
          paste0("clean_",max.gap,"_locs_AD_rouzic23.Rdata")))

source(here::here("R","problem_trips.R"))

##Interpolation
loc.clean.interp.rouzic23<-interpol_pastecs(data=cleanLocs.rouzic23,
                                           time.int=reso,
                                           colony=colo_coord_rouzic) %>% 
                         left_join(.,serol.rouzic, by=c("id"))


 save(loc.clean.interp.rouzic23,
     file=here::here("data","NewlyCreatedData",paste0("loc_clean_interp",max.gap,"_rouzic23.RData")))

SPDF <-st_as_sf(loc.clean.interp.rouzic23,
         coords = c("long", "lat"), 
        crs = 4326)

 st_write(SPDF,
           dsn=here("data","NewlyCreatedData","shapefiles"),
          layer= paste0("clean_",max.gap,"_locs_AD_rouzic23"),
          driver="ESRI Shapefile",
          delete_layer =T)
 
#generate maps for each individuals based on interpolated data
#  ids<-unique(loc.clean.interp.rouzic23$id)
# #
#  for (a in 1:length(ids)){
# 
# #     #extract data for 1 ind
#      indi<-loc.clean.interp.rouzic23 %>%
#         dplyr::filter(id==ids[a])
# 
#      #get sf layer for 1 ind
#     indisf<-loc.clean.interp.rouzic23 %>%
#         dplyr::filter(id==ids[a]) %>%
#         st_as_sf(.,coords=c("long","lat")) %>%
#         st_set_crs(st_crs(4326))
# 
#     #assign shape according to life stage
#     shap<-ifelse(indi$stage[1]=="juvenile",17,16)
# 
# 
#     #create external files
#     png(here::here("outputs","Trips by individuals",
#                    paste0(ids[a],"_interp_trips.png")),
#         width=5000,height=3000,res=400)
# 
#     #plot of maximal distance to the colony
#    distrawind<-ggplot(indi,aes(x=datetime,y=distmax))+
#     geom_point(aes(color=datetime),shape=shap,show.legend=F) +
#     labs(x="Date",y="Maximal distance (km)")
# 
#    #Map of all locations for 1 ind
#    map<-ggplot()+
#        geom_sf(data=indisf, aes(color=datetime),
#                shape=shap,
#                show.legend=F) +
#        geom_sf(data=shp,fill="grey90") +
#        coord_sf(xlim=c(floor(st_bbox(indisf)[1]),
#                        ceiling(st_bbox(indisf)[3])),
#                 ylim=c(floor(st_bbox(indisf)[2]),
#                        ceiling(st_bbox(indisf)[4]))) +
#   theme(panel.background = element_rect(fill = "lightblue"))
# 
#    grid.arrange(distrawind, map,ncol=2,
#               top = grid::textGrob(paste0("Interp data ",ids[a]),
#                                    gp=gpar(fontsize=14,font=3)))
#   dev.off()
# 
# }

```
