---
title: "1-Data processing"
author: "Aurore Ponchon"
format: html
editor: visual
---

```{r}
#| echo: false
#| include: true
#| warning: false
#| message: false


library(move)
library(here)
library(sf)
library(tidyverse)
library(lubridate)
library(conflicted)
library(readxl)
library(grid)
library(gridExtra)
library(sf)
library(rnaturalearth)

source(here::here("R","trip_functions.R"))
```

# 1-Importing tracking data

Data deposited on movebank.com. An account is required to access and download the data.

```{r}
#| echo: false
#| include: true
#| warning: false
#| message: false


#Import data directly from movebank
# loginStored <- movebankLogin(username="AurorePonchon",
#                                password="***")
# 
# gannets<-searchMovebankStudies(x="gannet",
#                                  login=loginStored)[1]

# metadata<-getMovebankStudy(study=gannets,
#                            login=loginStored)
# getMovebankSensorsAttributes(study=gannets,
#                              login=loginStored)
# reftable<-getMovebankReferenceTable(study=gannets,
#                                     login=loginStored)

# #raw data from 2022 and 2023 are downloaded and saved
# move.rouzic <- getMovebankData(study=gannets,
#                                 login=loginStored,
#                               removeDuplicatedTimestamps=TRUE)
# save(move.rouzic,
#       file=here("data","RawData","gannets_rouzic_raw_2022_2023.Rdata"))


#raw data from 2022 nd 2023 
load(here("data","RawData","gannets_rouzic_raw_2022_2023.Rdata"))

 #format raw data and select only useful columns
all_rouzic<-move.rouzic@data %>% 
    dplyr::select(tag_local_identifier,
                  timestamp,
                  location_long,
                  location_lat,
                  barometric_height, 
                  ground_speed) %>% 
    rename(id=tag_local_identifier,
           datetime=timestamp,
           long=location_long,
           lat=location_lat,
           alt=barometric_height,
           speed=ground_speed)  %>% 
    arrange(id,datetime) %>% 
    mutate(id=as.factor(id),
           site=as.factor("Rouzic"),
           year=year(datetime))


#convert formatted raw data into shapefile
# rouziclayer<- st_as_sf(all_rouzic,
#                        coords = c("long", "lat"), 
#                        crs = 4326)
# st_write(rouziclayer,
#          dsn=here("data","NewlyCreatedData","shapefiles"),
#          layer="all_raw_gannets_rouzic_2022_2023",
#          driver="ESRI Shapefile",
#          delete_layer =T,
#           delete_dsn = T)

```

# Importing individual data

Individual data (sex, date of equipment, breeding stage, period, serological status etc...) are stored in a separate file

```{r}
serol.rouzic<-read_excel(here("data","RawData","data_Rouzic_2023_sexage_AIV.xlsx"),
              sheet="capture_data",col_names = T) %>% 
      dplyr::select(GSM,
                    stage,
                    date,
                    sexe,
                    age_chick,
                    AIV,
                    body_weight,
                    tarsus,
                    wing,
                    culmen) %>% 
       rename(id=GSM, 
              sex=sexe,
              sero=AIV,
              period=date) %>% 
       mutate(id=as.factor(id),
              stage=as.factor(stage),
              sex=as.factor(sex),
              period=as.Date(period,format="%F"),
              deploy=ifelse(period < as.Date("2023-08-01",
                                        format="%F"),
                            1,
                            2))  %>% 
     dplyr::filter(id!="." & id!="233661.1") %>% 
    droplevels()
    
```

# 3- Classifying individual at-sea trips

Locations are classified in consecutive trips at-sea when the birds leave the colony further than 1km away from the colony and for more than 1h. Some preliminary plots are used to identify the duration of time gaps.

```{r}
all_rouzic23<-all_rouzic %>% 
              dplyr::filter(year==2023) %>% 
              droplevels() %>% 
              left_join(., serol.rouzic, by="id",keep=F)
  
remove(all_rouzic)  

ids<-unique(levels(all_rouzic23$id))
colo.ind<-tim.colors(length(ids))

#create new object with distance and time differences
raw_rouzic_2023<-NULL

#loop through each individual 
for (i in 1:length(ids)){
    ind.loc<-all_rouzic23 %>% 
        dplyr::filter(id==ids[i]) %>% 
        mutate(distmax=0,
               distadj=0,
               totalpath=0,
               difftimemin=0,
               travelNb=0,
               onlandNb=0)
  
    print(c("Processing individual ", i))
      
  #loop through each line to calculate distance and time difference with previous location
for (l in 2:nrow(ind.loc)){
    
ind.loc[l,"distadj"]<-rdist.earth(ind.loc[l,c("long","lat")],
                                   ind.loc[l-1,c("long","lat")], 
                                  miles = F)
ind.loc[l,"difftimemin"]<-difftime(ind.loc[l,"datetime"],
                                   ind.loc[l-1,"datetime"], 
                                   units="mins")
ind.loc[l,"distmax"]<-rdist.earth(ind.loc[l,c("long","lat")],
                                   colo_coord_rouzic[c(1,2)], 
                                  miles = F)
}

#define periods on land and at sea   
ind.loc<-define_trips(data=ind.loc,
                      dist.min=dist.threshold)
raw_rouzic_2023<-rbind(raw_rouzic_2023,ind.loc)    
} #end of loop on individuals


raw_rouzic_2023$trip.id<-paste(raw_rouzic_2023$id,raw_rouzic_2023$travelNb,sep=".")
raw_rouzic_2023<-add_missing_return(raw_rouzic_2023,
                                  time.int=reso,colony=colo_coord_rouzic)

save(raw_rouzic_2023,file=here("data","NewlyCreatedData","rawLocRouzic2023.Rdata"))

load(here("data","NewlyCreatedData","rawLocRouzic2023.Rdata"))

#plot distance to the colony over time
distraw<-ggplot(raw_rouzic_2023,aes(x=datetime,y=distmax),group=id)+
    geom_point(aes(color=id),show.legend=F) +
    scale_y_continuous(limits=c(0,200)) +
    scale_color_manual(values=colo.ind)+
    labs(x="Date",y="Maximal distance to colony (km)")

#plot time gaps
diffraw<-ggplot(raw_rouzic_2023,aes(x=datetime,y=difftimemin),group=id)+
    geom_point(aes(color=id),show.legend=F) +
    scale_y_continuous(limits=c(0,3600)) +
    scale_color_manual(values=colo.ind)+
    labs(x="Date",y="Time gaps between 2 locations (in min)")

grid.arrange(distraw,diffraw, ncol=2,
              top = grid::textGrob("Raw data",gp=gpar(fontsize=16,font=3)))

#get overview of trips for each individual

target<-c("France","United Kingdom","Spain","Morocco","Belgium",
          "Netherlands")

#contours of countries
shp<-ne_countries(scale = 10, returnclass = 'sf') %>%
   dplyr::filter(name %in% target) 

for (a in 1:length(ids)){
    
    #extract data for 1 ind
     indi<-raw_rouzic_2023 %>% 
        dplyr::filter(id==ids[a])
    
     #get sf layer for 1 ind
    indisf<-raw_rouzic_2023 %>% 
        dplyr::filter(id==ids[a]) %>% 
        st_as_sf(.,coords=c("long","lat")) %>% 
        st_set_crs(st_crs(4326))
    
#assign shape according to life stage
    shap<-ifelse(indi$stage[1]=="juvenile",17,16)

    #create external files
    png(here::here("outputs","Raw trips by individuals",
                   paste0("Raw_trips",ids[a],".png")),
        width=5000,height=3000,res=400)

    #plot of maximal distance to the colony    
   distrawind<-ggplot(indi,aes(x=datetime,y=distmax))+
    geom_point(aes(color=datetime),shape=shap,show.legend=F) +
    labs(x="Date",y="Maximal distance (km)")
 
   #plot of time gaps
   diffrawind<-ggplot(indi,aes(x=datetime,y=difftimemin,
                               group=stage))+
    geom_point(aes(color=datetime),shape=shap,
               show.legend=F) +
    labs(x="Date",y="Time gap (in min)")
   
   #Map of all locations for 1 ind
   map<-ggplot()+
       geom_sf(data=indisf, aes(color=datetime),
               shape=shap,
               show.legend=F) +
       geom_sf(data=shp,fill="grey90") +
       
       coord_sf(xlim=c(floor(st_bbox(indisf)[1]),
                       ceiling(st_bbox(indisf)[3])),
                ylim=c(floor(st_bbox(indisf)[2]),
                       ceiling(st_bbox(indisf)[4]))) +
  theme(panel.background = element_rect(fill = "lightblue"))

   grid.arrange(distrawind,diffrawind, map,ncol=3,
              top = grid::textGrob(paste0("Raw data ",ids[a]),
                                   gp=gpar(fontsize=14,font=3)))
  dev.off()
  
}

```

# 4- Summary of raw at-sea trips

Calculation of at-sea trip characteristics and periods on land

```{r}


#summarize all trips and periods on land by individuals
raw_trips_rouzic23<-trips_summary_ind(raw_rouzic_2023,colony=colo_coord_rouzic)
raw_land_rouzic23<-land_summary_ind(raw_rouzic_2023,colony=colo_coord_rouzic)


summy.trips<-raw_trips_rouzic23 %>% 
    group_by(id) %>% 
    summarize(Nbtrip=n(),
              Mindurh=min(TripDurh),
              maxDiffh=max(maxDiffTimeh),
              minLocs=min(nlocs))

#details individual trips
# kbl(summy.trips,caption="Summary of raw trips") %>%
#   kable_styling(bootstrap_options = c("striped", "hover"))


summy.land<-raw_land_rouzic23 %>% 
    group_by(id) %>% 
    summarize(Nbland=n(),
              Mindurh=min(LandDurh),
              maxDiffh=max(maxDiffTimeh),
              minLocs=min(nlocs))


##plot trip and land periods
dist<-ggplot(raw_trips_rouzic23,aes(x=DateEnd,y=Distmaxkm,group=id))+
    geom_point(aes(color=id, shape=stage),show.legend=F) +
 #   scale_shape_manual(values=c(16,17)) +
    scale_y_continuous(limits=c(0,500)) +
    scale_color_manual(values=colo.ind) +
    labs(x="Date",y="Maximal distance to colony (km)")

dur<-ggplot(raw_trips_rouzic23,aes(x=DateEnd,y=TripDurh ,group=id))+
    geom_point(aes(color=id,shape=stage),show.legend=F) +
  #  scale_y_continuous(limits=c(0,10)) +
    scale_color_manual(values=colo.ind)+
    labs(x="Date",y="Trip duration (h")

correl<-ggplot(raw_trips_rouzic23,aes(x=log(TripDurh),y=log(Distmaxkm)))+
    geom_point(aes(color=DateEnd, shape=stage), show.legend=F) +
    scale_colour_datetime(low=viridis(100)[1],high=viridis(100)[100]) +
    labs(x="log Trip duration (h)",y="log Maximal distance to colony (km)")

land<-ggplot(raw_land_rouzic23,aes(x=DateEnd,y=LandDurh ,group=id))+
    geom_point(aes(color=id,shape=stage), show.legend=F) +
  #  scale_y_continuous(limits=c(0,10)) +
    scale_color_manual(values=colo.ind)+
    labs(x="Date",y="Periods on land (h)")

grid.arrange(dist,dur,correl, land, ncol=2, nrow=2,
              top = textGrob("Raw trip features",gp=gpar(fontsize=16,font=3)))

```
