---
title: "2-Trip characteristics"
author: "Aurore Ponchon"
format: html
editor: visual
---

## Trip summary with gaps \<6h

```{r}
#| echo: false
#| include: true
#| warning: false
#| message: false

library(here)
library(tidyverse)
library(gridExtra)
library(nlme)
library(emmeans)
library(DHARMa)
library(kableExtra)
library(plotrix)

source(here("R","trip_functions.R"))

mycol<-turbo(2,begin=0.25, end=0.7)

#trips
load(here::here("data","NewlyCreatedData","loc_clean_interp6_rouzic23.RData"))
trips6<-trips_summary_ind(loc.clean.interp.rouzic23,
                                  colony=colo_coord_rouzic) %>% 
    dplyr::mutate(sero=as.factor(sero))

#time in colony
load(here("data","NewlyCreatedData","rawLocRouzic2023.Rdata"))
land<-raw_rouzic_2023 %>% 
    land_summary_ind(.,colony=colo_coord_rouzic) %>% 
    dplyr::filter(LandDurh > 1 & 
                    LandDurh < 100 & 
                      stage=="adult") %>% 
     dplyr::mutate(sero=as.factor(sero))


library(FactoMineR)
library(corrr)
library(corrplot)
library(factoextra)
library(readxl)
library(ggcorrplot)


serol.rouzic<-get_serol_data_2023()   
             
#summary of all trips (26 adults)
trips<-trips6 %>% 
       mutate(day=as.numeric(as.Date(DateEnd) - as.Date("2023-07-02"))) %>% 
       droplevels() %>% 
              left_join(., serol.rouzic, by=c("id","sex","sero",
                                              "deploy","stage"),keep=F)


lands<-land %>% 
      mutate(day=as.numeric(as.Date(DateEnd) - as.Date("2023-07-02"))) %>% 
      droplevels() %>% 
      left_join(., serol.rouzic, by=c("id","sex","sero","deploy","stage"),keep=F)

#get the 4 numerical variables which are used as proxies of body condition 
mat<-trips %>% 
    dplyr::distinct(id,body_weight,wing, culmen,tarsus, .keep_all = F) %>% 
    dplyr::select(-id) %>% 
    scale(.)

#check correlation between variables
corr_matrix <- cor(mat)
corrplot(corr_matrix, method="number")

#pca
data.pca <- princomp(mat)
summary(data.pca)
fviz_pca_var(data.pca, col.var = "cos2",
            gradient.cols = c("black", "orange", "green"),
            repel = TRUE)

#extract pca results
cond<-get_pca_ind(data.pca)

#extract the axis 1 values
IDS<-trips %>% 
      dplyr::distinct(id,body_weight,wing, culmen, .keep_all = F) %>% 
    dplyr::mutate(body_cond=cond$coord[,1])

#merge body condition index with rest of data
trips.cond<-trips %>% 
    left_join(.,IDS,by=c("id","body_weight","culmen","wing"))
lands.cond<-lands %>% 
    left_join(.,IDS,by=c("id","body_weight","culmen","wing"))


#Table summary
sum.ids<-trips6 %>% 
    summarize(Ids=length(unique(id)),
              Trips=length(unique(trip.id)),
              Debut=min(DateEnd),
              Fin=max(DateEnd))
  
kbl(sum.ids,caption="Sample size for trips") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))


sum.id<- trips6 %>% 
    group_by(id) %>% 
    summarize(across(c(Distmaxkm, TripDurh, TotalPathkm), 
                     ~ mean(.x, na.rm = TRUE))) %>% 
    as.data.frame()

sum.trips1<- sum.id %>% 
    summarize(across(where(is.numeric), 
                     list(Mean=mean,
                          SE=plotrix::std.error,
                          Min=min,
                          Max=max))) %>% 
    mutate(across(where(is.numeric), round,digits=1))


sum.trips<-sum.trips1 %>% 
           dplyr::mutate(MeanDist=paste(Distmaxkm_Mean,Distmaxkm_SE,sep=" ± "),
                         DistRange=paste(Distmaxkm_Min,Distmaxkm_Max,sep=" - "),
                         MeanDur=paste(TripDurh_Mean,TripDurh_SE,sep=" ± "),
                         DurRange=paste(TripDurh_Min,TripDurh_Max,sep=" - "),     
                         MeanPath=paste(TotalPathkm_Mean,TotalPathkm_SE,sep=" ± "),
                      PathRange=paste(TotalPathkm_Min,TotalPathkm_Max,sep=" - ")) %>%
        dplyr::select(-c(1:12))

lands.ids<-land %>% 
    summarize(Ids=length(unique(id)),
              Land=length(unique(onland.id)),
              Debut=min(DateEnd),
              Fin=max(DateEnd))

kbl(lands.ids,caption="Sample size for periods on land") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

land.id<- land %>% 
    group_by(id) %>% 
    summarize(Time= mean(LandDurh, na.rm=T)) %>% 
    as.data.frame()

sum.lands<- land.id %>% 
    summarize(across(where(is.numeric), 
                     list(mean=mean,
                          se=plotrix::std.error))) %>% 
    mutate(across(where(is.numeric), round,digits=1)) 

kbl(sum.trips,caption="Summary of trip characteristics - Gaps <6h",
    col.names=c("Maximal distance (km)",
                "Maximal distance (km)",
                "Trip Duration (h)",
                "Trip Duration (h)",
                "Total path (km)",
                "Total path (km)")) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

kbl(sum.lands,caption="Summary of nest attendance",
    col.names=c("DurationMean","DurationSE")) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

```

## Trip summary with gaps \<3h

```{r}
#| echo: false
#| include: true
#| warning: false
#| message: false

#trips
load(here::here("data","NewlyCreatedData","loc_clean_interp3_rouzic23.RData"))
trips3<-trips_summary_ind(loc.clean.interp.rouzic23,
                                  colony=colo_coord_rouzic) %>%
    dplyr::mutate(sero=as.factor(sero))

#Table summary
sum.ids<-trips3 %>%
    summarize(Ids=length(unique(id)),
              Trips=length(unique(trip.id)),
              Debut=min(DateEnd),
              Fin=max(DateEnd))

kbl(sum.ids,caption="Sample size for trips") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))


sum.id<- trips3 %>%
    group_by(id) %>%
    summarize(across(c(Distmaxkm, TripDurh, TotalPathkm),
                     ~ mean(.x, na.rm = TRUE))) %>%
    as.data.frame()

sum.trips1<- sum.id %>% 
    summarize(across(where(is.numeric), 
                     list(Mean=mean,
                          SE=plotrix::std.error,
                          Min=min,
                          Max=max))) %>% 
    mutate(across(where(is.numeric), round,digits=1))


sum.trips<-sum.trips1 %>% 
           dplyr::mutate(MeanDist=paste(Distmaxkm_Mean,Distmaxkm_SE,sep=" ± "),
                         DistRange=paste(Distmaxkm_Min,Distmaxkm_Max,sep=" - "),
                         MeanDur=paste(TripDurh_Mean,TripDurh_SE,sep=" ± "),
                         DurRange=paste(TripDurh_Min,TripDurh_Max,sep=" - "),     
                         MeanPath=paste(TotalPathkm_Mean,TotalPathkm_SE,sep=" ± "),
                      PathRange=paste(TotalPathkm_Min,TotalPathkm_Max,sep=" - ")) %>%
        dplyr::select(-c(1:12))

kbl(sum.trips,caption="Summary of trip characteristics - Gaps <3h",
    col.names=c("Maximal distance (km)",
                "Maximal distance (km)",
                "Trip Duration (h)",
                "Trip Duration (h)",
                "Total path (km)",
                "Total path (km)")) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))


```

# Comparisons between gaps \<3h and \<6h

```{r}
#| echo: false
#| include: true
#| warning: false
#| message: false


trips6$Group<-"H6"
trips3$Group<-"H3"
    
trips<-rbind(trips6,trips3)


 #maximal distance
dist<-ggplot(trips,aes(x=Group,y=Distmaxkm, fill=Group,
                                       colour=Group) ) +
  geom_boxplot(outlier.alpha = 0, alpha=0.3, show.legend = FALSE)+
  geom_jitter(width=0.1, show.legend = FALSE)+
  scale_y_continuous(limits=c(0,400),breaks=seq(0,400,100),expand=c(0.01,0.01))+
 labs(y="Maximal distance to the colony (km)",tag="a)",x="") +
  scale_fill_manual(values=mycol) +
  scale_colour_manual(values=mycol)  +
  theme(text = element_text(size=10),
        axis.text = element_text(size=10),
        legend.text = element_text(size=10),
        axis.title.x=element_blank())

#trip duration
dur<-ggplot(trips,aes(x=Group,y=TripDurh, 
                            fill=Group, colour=Group) )+
  geom_jitter(width=0.1, show.legend = FALSE)+
  geom_boxplot(outlier.alpha = 0, alpha=0.3, show.legend = FALSE)+
  labs(y="Trip duration (hrs)", x="",tag="b)") +
  scale_y_continuous(limits=c(0,200),expand=c(0.01,0.01))+
  scale_fill_manual(values=mycol) +
scale_colour_manual(values=mycol)+
  theme(text = element_text(size=10),
        axis.text = element_text(size=10),
        legend.text = element_text(size=10),
        axis.title.x=element_blank())

#total distance travelled
tot<-ggplot(trips,aes(x=Group,y=TotalPathkm, 
                                      fill=Group, colour=Group) )+
  geom_jitter(width=0.1, show.legend = FALSE)+
  geom_boxplot(outlier.alpha = 0, alpha=0.3, show.legend = FALSE)+
  labs(y="Total distance travelled (km)",x="",tag="c)") +
  scale_y_continuous(limits=c(0,1200),expand=c(0.01,0.01))+
  scale_fill_manual(values=mycol) +
  scale_colour_manual(values=mycol)+
  theme(text = element_text(size=10),
        axis.text = element_text(size=10),
        legend.text = element_text(size=10),
        axis.title.x=element_blank())

#nest attendance
att<-ggplot(land,aes(x=site,y=LandDurh))+
  geom_jitter(width=0.1,show.legend=F)+
  geom_boxplot(outlier.alpha = 0, alpha=0.3, show.legend = FALSE)+
  labs(y="Nest attendance (hrs)",x="",tag="d)") +
#  scale_y_continuous(limits=c(0,120),expand=c(0.01,0.01))+
  scale_fill_manual(values=mycol)  +
  scale_colour_manual(values=mycol)+
  theme(text = element_text(size=10),
        axis.text = element_text(size=10),
        axis.title.x=element_blank())

 # tiff("Figures/Figure 2 - Stats trip characteristics.tif",compression="lzw",
 #      res=600,width=6500,height=3500)
grid.arrange(arrangeGrob(dist,dur,tot,att,
                         ncol=2,nrow=2))



 #maximal distance
distb<-ggplot(trips,aes(x=DateStart,y=Distmaxkm ,group=sero)) +
   geom_point(aes(color=sero,shape=sero),show.legend = F)+

    facet_wrap(Group~., nrow=2) +
 labs(y="Maximal distance to the colony (km)",tag="a)",x="") +
  scale_fill_manual(values=mycol) +
  scale_colour_manual(values=mycol)  

 #max duration
durb<-ggplot(trips,aes(x=DateStart,y=TripDurh ,group=sero)) +
   geom_point(aes(color=sero,shape=sero),show.legend = F)+
    facet_wrap(Group~., nrow=2) +
 labs(y="Trip duration (hrs)",tag="b)",x="") +
  scale_fill_manual(values=mycol) +
  scale_colour_manual(values=mycol)  

 #total distance
totb<-ggplot(trips,aes(x=DateStart,y=TotalPathkm ,
                        group=sero)) +
   geom_point(aes(color=sero,shape=sero),show.legend = F)+
    facet_wrap(Group~., nrow=2) +
 labs(y="Total distance travelled (km)",tag="c)",x="") +
  scale_fill_manual(values=mycol) +
  scale_colour_manual(values=mycol)  

grid.arrange(distb,durb,totb,
             ncol=3)
```

# Comparison between infected and non-infected individuals

```{r}
#| echo: false
#| include: true
#| warning: false
#| message: false

 #maximal distance
trips6$sero<-as.factor(trips6$sero)

dist.sero<-ggplot(trips6,aes(x=sero,y=Distmaxkm, fill=sero,
                                       colour=sero) ) +
    geom_violin(trim=FALSE, alpha=0.2)+
  geom_boxplot(outlier.alpha = 0, alpha=0.3, show.legend = FALSE)+
  geom_jitter(width=0.2) +
  scale_y_continuous(limits=c(0,400),breaks=seq(0,400,100),expand=c(0.01,0.01))+
 labs(y="Maximal distance to the colony (km)",tag="a)",x="") +
  scale_fill_manual(values=mycol) +
  scale_colour_manual(values=mycol)  +
  theme(text = element_text(size=10),
        axis.text = element_text(size=10),
        legend.text = element_text(size=10),
        axis.title.x=element_blank())


#trip duration
dur.sero<-ggplot(trips6,aes(x=sero,y=TripDurh, 
                            fill=sero, colour=sero) )+
      geom_violin(trim=FALSE, alpha=0.2)+
  geom_jitter(width=0.2)+
  geom_boxplot(outlier.alpha = 0, alpha=0.3, show.legend = FALSE)+
  labs(y="Trip duration (hrs)", x="",tag="b)") +
  scale_y_continuous(limits=c(0,200),expand=c(0.01,0.01))+
  scale_fill_manual(values=mycol) +
scale_colour_manual(values=mycol)+
  theme(text = element_text(size=10),
        axis.text = element_text(size=10),
        legend.text = element_text(size=10),
        axis.title.x=element_blank())

#total distance travelled
tot.sero<-ggplot(trips6,aes(x=sero,y=TotalPathkm, 
                                      fill=sero, colour=sero) )+
      geom_violin(trim=FALSE, alpha=0.2)+
  geom_jitter(width=0.2)+
  geom_boxplot(outlier.alpha = 0, alpha=0.3, show.legend = FALSE)+
  labs(y="Total distance travelled (km)",x="",tag="c)") +
  scale_y_continuous(limits=c(0,1200),expand=c(0.01,0.01))+
  scale_fill_manual(values=mycol, name="Infected") +
  scale_colour_manual(values=mycol, name="Infected")+
  theme(text = element_text(size=10),
        axis.text = element_text(size=10),
        legend.text = element_text(size=10),
        axis.title.x=element_blank())

#nest attendance
land<-land %>% 
    dplyr::mutate(Sero=as.factor(sero))

att.sero<-ggplot(land,aes(x=Sero,y=LandDurh,
                     fill=Sero, colour=Sero))+
      geom_violin(trim=FALSE, alpha=0.2)+
  geom_jitter(width=0.2) +
  geom_boxplot(outlier.alpha = 0, alpha=0.3, show.legend = FALSE) +
  labs(y="Nest attendance (hrs)",x="",tag="d)") +
  scale_y_continuous(limits=c(0,60),expand=c(0.01,0.01))+
  scale_fill_manual(values=mycol)  +
  scale_colour_manual(values=mycol)+
  theme(legend.position = "none",
      text = element_text(size=10),
        axis.text = element_text(size=10),
        axis.title.x=element_blank())

mylegend<-g_legend(tot.sero)

 # tiff("Figures/Figure 2 - Stats trip characteristics.tif",compression="lzw",
 #      res=600,width=6500,height=3500)

tiff(here("outputs","Trip_characteristics_infected_2023.tif"),
     width=7000,
     height=4000,
     res=600,
     compression="lzw")
grid.arrange(arrangeGrob(dist.sero+ theme(legend.position="none"),
                         dur.sero+ theme(legend.position="none"),
                         tot.sero+ theme(legend.position="none"),
                         att.sero+ theme(legend.position="none") ,
                                     ncol=2),
               mylegend, ncol=2,widths=c(1,0.1)) 

    dev.off()  

```

# Comparison between sex

```{r}
#| echo: false
#| include: true
#| warning: false
#| message: false

mycol2<-magma(n=2,begin=0.3,end=0.7)

trips6$sex<-as.factor(trips6$sex)

#maximal distance
dist.sex<-ggplot(trips6,aes(x=sex,y=Distmaxkm, fill=sex,
                                       colour=sex) ) +
    geom_violin(trim=FALSE, alpha=0.2)+
  geom_boxplot(outlier.alpha = 0, alpha=0.3, show.legend = FALSE)+
  geom_jitter(width=0.2) +
  scale_y_continuous(limits=c(0,400),breaks=seq(0,400,100),expand=c(0.01,0.01))+
 labs(y="Maximal distance to the colony (km)",tag="a)",x="") +
  scale_fill_manual(values=mycol2) +
  scale_colour_manual(values=mycol2)  +
  theme(text = element_text(size=10),
        axis.text = element_text(size=10),
        legend.text = element_text(size=10),
        axis.title.x=element_blank())


#trip duration
dur.sex<-ggplot(trips6,aes(x=sex,y=TripDurh, 
                            fill=sex, colour=sex) )+
      geom_violin(trim=FALSE, alpha=0.2)+
  geom_jitter(width=0.2)+
  geom_boxplot(outlier.alpha = 0, alpha=0.3, show.legend = FALSE)+
  labs(y="Trip duration (hrs)", x="",tag="b)") +
  scale_y_continuous(limits=c(0,200),expand=c(0.01,0.01))+
  scale_fill_manual(values=mycol2) +
scale_colour_manual(values=mycol2)+
  theme(text = element_text(size=10),
        axis.text = element_text(size=10),
        legend.text = element_text(size=10),
        axis.title.x=element_blank())

#total distance travelled
tot.sex<-ggplot(trips6,aes(x=sex,y=TotalPathkm, 
                                      fill=sex, colour=sex) )+
      geom_violin(trim=FALSE, alpha=0.2)+
  geom_jitter(width=0.2)+
  geom_boxplot(outlier.alpha = 0, alpha=0.3, show.legend = FALSE)+
  labs(y="Total distance travelled (km)",x="",tag="c)") +
  scale_y_continuous(limits=c(0,1200),expand=c(0.01,0.01))+
  scale_fill_manual(values=mycol2, name="Sex") +
  scale_colour_manual(values=mycol2, name="Sex")+
  theme(text = element_text(size=10),
        axis.text = element_text(size=10),
        legend.text = element_text(size=10),
        axis.title.x=element_blank())

#nest attendance
land<-land %>% 
    dplyr::mutate(sero=as.factor(sero))

att.sex<-ggplot(land,aes(x=sex,y=LandDurh,
                     fill=sex, colour=sex))+
      geom_violin(trim=FALSE, alpha=0.2)+
  geom_jitter(width=0.2) +
  geom_boxplot(outlier.alpha = 0, alpha=0.3, show.legend = FALSE) +
  labs(y="Nest attendance (hrs)",x="",tag="d)") +
  scale_y_continuous(limits=c(0,60),expand=c(0.01,0.01))+
  scale_fill_manual(values=mycol2)  +
  scale_colour_manual(values=mycol2)+
  theme(legend.position = "none",
      text = element_text(size=10),
        axis.text = element_text(size=10),
        axis.title.x=element_blank())

mylegendsex<-g_legend(tot.sex)

 # tiff("Figures/Figure 2 - Stats trip characteristics.tif",compression="lzw",
 #      res=600,width=6500,height=3500)

tiff(here("outputs","Trip_characteristics_sex_2023.tif"),
     width=7000,
     height=4000,
     res=600,
     compression="lzw")
grid.arrange(arrangeGrob(dist.sex+ theme(legend.position="none"),
                         dur.sex+ theme(legend.position="none"),
                         tot.sex+ theme(legend.position="none"),
                         att.sex+ theme(legend.position="none") ,
                                     ncol=2),
               mylegendsex, ncol=2,widths=c(1,0.1)) 

    dev.off()  

```

# Comparison between deployment 1 & 2

```{r}
#| echo: false
#| include: true
#| warning: false
#| message: false

trips6$deploy<-as.factor(trips6$deploy)

dist.deploy<-ggplot(trips6,aes(x=deploy,y=Distmaxkm, fill=deploy,
                                       colour=deploy) ) +
    geom_violin(trim=FALSE, alpha=0.2)+
  geom_boxplot(outlier.alpha = 0, alpha=0.3, show.legend = FALSE)+
  geom_jitter(width=0.1, show.legend = FALSE) +
  scale_y_continuous(limits=c(0,400),breaks=seq(0,400,100),expand=c(0.01,0.01))+
 labs(y="Maximal distance to the colony (km)",tag="a)",x="") +
  scale_fill_manual(values=mycol) +
  scale_colour_manual(values=mycol)  +
  theme(text = element_text(size=10),
        axis.text = element_text(size=10),
        legend.text = element_text(size=10),
        axis.title.x=element_blank())


#trip duration
dur.deploy<-ggplot(trips6,aes(x=deploy,y=TripDurh, 
                            fill=deploy, colour=deploy) )+
      geom_violin(trim=FALSE, alpha=0.2)+
  geom_jitter(width=0.1, show.legend = FALSE)+
  geom_boxplot(outlier.alpha = 0, alpha=0.3, show.legend = FALSE)+
  labs(y="Trip duration (hrs)", x="",tag="b)") +
  scale_y_continuous(limits=c(0,200),expand=c(0.01,0.01))+
  scale_fill_manual(values=mycol) +
scale_colour_manual(values=mycol)+
  theme(text = element_text(size=10),
        axis.text = element_text(size=10),
        legend.text = element_text(size=10),
        axis.title.x=element_blank())

#total distance travelled
tot.deploy<-ggplot(trips6,aes(x=deploy,y=TotalPathkm, 
                                      fill=deploy, colour=deploy) )+
      geom_violin(trim=FALSE, alpha=0.2)+
  geom_jitter(width=0.1, show.legend = FALSE)+
  geom_boxplot(outlier.alpha = 0, alpha=0.3, show.legend = FALSE)+
  labs(y="Total distance travelled (km)",x="",tag="c)") +
  scale_y_continuous(limits=c(0,1200),expand=c(0.01,0.01))+
  scale_fill_manual(values=mycol) +
  scale_colour_manual(values=mycol)+
  theme(text = element_text(size=10),
        axis.text = element_text(size=10),
        legend.text = element_text(size=10),
        axis.title.x=element_blank())

#nest attendance
land<-land %>% 
    dplyr::mutate(deploy=as.factor(deploy))

att.deploy<-ggplot(land,aes(x=deploy,y=LandDurh,
                     fill=deploy, colour=deploy))+
      geom_violin(trim=FALSE, alpha=0.2)+
  geom_jitter(width=0.1,show.legend=F) +
  geom_boxplot(outlier.alpha = 0, alpha=0.3, show.legend = FALSE) +
  labs(y="Nest attendance (hrs)",x="",tag="d)") +
#  scale_y_continuous(limits=c(0,120),expand=c(0.01,0.01))+
  scale_fill_manual(values=mycol)  +
  scale_colour_manual(values=mycol)+
  theme(text = element_text(size=10),
        axis.text = element_text(size=10),
        axis.title.x=element_blank())

 # tiff("Figures/Figure 2 - Stats trip characteristics.tif",compression="lzw",
 #      res=600,width=6500,height=3500)
grid.arrange(arrangeGrob(dist.deploy,dur.deploy,tot.deploy,att.deploy,
                         ncol=2,nrow=2))

```

# Statistical tests with serology only

```{r}
#| echo: false
#| include: true
#| warning: false
#| message: false

source(here::here("R","stats_functions.R"))
var<-c('day','sero')

#summary of all trips (26 adults)
trips<-trips6 %>% 
       mutate(deploy=as.factor(deploy),
           day=as.numeric(as.Date(DateEnd) - as.Date("2023-07-02")))


lands<-land %>% 
      mutate(deploy=as.factor(deploy),
           day=as.numeric(as.Date(DateEnd) - as.Date("2023-07-02")))



#daysx<-c("03/07","20/08","30/08","09/09","19/09","29/09","20/10")
newdat<-expand.grid(site=unique(trips$site),
                    day=seq(min(trips$day),
                            max(trips$day),1),
                    sero=unique(trips$sero))

newdatland<-expand.grid(site=unique(lands$site),
                    day=seq(min(lands$day),
                            max(lands$day),1),
                    sero=unique(lands$sero))

#stats for maximal distance
ggplot(trips, aes(x=day, y=Distmaxkm, 
                  colour=id)) +
    geom_point()+
    geom_smooth(method="lm",alpha=0.2)+
    ggtitle("Maximal distance")


glm.dist<-lme(sqrt(Distmaxkm)~ day + sero,
              random= ~1| id,
              data=trips)

summary(glm.dist)
anova(glm.dist)

qqnorm(resid(glm.dist))
qqline(resid(glm.dist))


#stats for trip duration
ggplot(trips, aes(x=day, y=TripDurh, 
                  colour=id)) +
    geom_point()+
    geom_smooth(method="lm",alpha=0.2)+
    ggtitle("Trip duration")


glm.dur<-lme(sqrt(TripDurh)~ day + sero,
              random= ~1| id,
              data=trips)

summary(glm.dur)
anova(glm.dur)

qqnorm(resid(glm.dur))
qqline(resid(glm.dur))


##stats for total path
ggplot(trips, aes(x=day, y=TotalPathkm, 
                  colour=id)) +
    geom_point()+
    geom_smooth(method="lm",alpha=0.2)+
    ggtitle("Total path")


glm.path<-lme(sqrt(TotalPathkm)~ day + sero,
              random= ~1| id,
              data=trips)

summary(glm.path)
anova(glm.path)

qqnorm(resid(glm.path))
qqline(resid(glm.path))

##stats for time on land
glm.land<-lme(sqrt(LandDurh)~ day + sero,
              random= ~1| id,
              data=lands)

summary(glm.land)
anova(glm.land)

qqnorm(resid(glm.land))
qqline(resid(glm.land))


#predict fitted values and CI

predist<-predict_lme_results(glm.dist,newdat,var)
preddur<-predict_lme_results(glm.dur,newdat,var)
predpath<-predict_lme_results(glm.path,newdat,var)
predland<-predict_lme_results(glm.land,newdatland,var)

#mycol<-viridis(n=2)

distlmmsero<-ggplot(trips, aes(x=day, y=Distmaxkm ,colour=sero)) + 
    geom_ribbon(data= data.frame(predist),
            aes(ymin= lower.CL^2, ymax= upper.CL^2,
                y= NULL,
                fill=sero),
           alpha=0.4,colour = NA) +
geom_point(alpha=0.6) +
 geom_line(data=predist,aes(x=day,y=emmean^2),lwd=1.5) +
     labs(y="Maximal distance to the colony (km)",tag="a)",x="") +
    scale_x_continuous(limits=c(0,120),
                     #  labels=daysx,
                       breaks=seq(0,120,20),
                       expand=c(0.001,0.001),
                     labels=c("02/07","22/07","11/08","31/08","20/09","10/10","30/10")) +
     scale_y_continuous(limits=c(0,400,100),
                        breaks=seq(0,400,100),
                        expand=c(0.001,0.001)) + 
 scale_colour_manual(values=mycol,name="Infected") +
 scale_fill_manual(values=mycol,name="Infected")  +
     theme_light()+
    theme(#legend.position="none",
          panel.grid.minor = element_blank()) 

durlmmsero<-ggplot(trips, aes(x=day, y=TripDurh ,colour=sero)) + 
    geom_ribbon(data= data.frame(preddur), 
            aes(ymin= lower.CL^2, ymax= upper.CL^2, 
                y= NULL,
                fill=sero), 
            alpha=0.4,colour = NA) +
geom_point(alpha=0.6) +
 geom_line(data=preddur,aes(x=day,y=emmean^2),lwd=1.5) +
     labs(y="Trip Duration (h)",tag="b)",x="") +
    scale_x_continuous(limits=c(0,120),
                     #  labels=daysx,
                       breaks=seq(0,120,20),
                       expand=c(0.001,0.001),
                     labels=c("02/07","22/07","11/08","31/08","20/09","10/10","30/10")) +
    scale_y_continuous(limits=c(0,200),
                       breaks=seq(0,200,50),
                       expand=c(0.001,0)) +
  scale_colour_manual(values=mycol,name="Infected") +
    scale_fill_manual(values=mycol,name="Infected")  +
    theme_light()+
    theme(#legend.position="none",
          panel.grid.minor = element_blank()) 
    

pathlmmsero<-ggplot(trips, aes(x=day, y=TotalPathkm ,colour=sero)) + 
    geom_ribbon(data= data.frame(predpath), 
            aes(ymin= lower.CL^2, ymax= upper.CL^2, 
                y= NULL,
                fill=sero), 
            alpha=0.4,colour = NA) +
geom_point(alpha=0.6) +
 geom_line(data=predpath,aes(x=day,y=emmean^2),lwd=1.5) +
     labs(y="Total distance travelled (km)",tag="c)",x="") +
    scale_x_continuous(limits=c(0,120),
                     #  labels=daysx,
                       breaks=seq(0,120,20),
                       expand=c(0.001,0.001),
                     labels=c("02/07","22/07","11/08","31/08","20/09","10/10","30/10")) +
    scale_y_continuous(limits=c(0,1200),
                       breaks=seq(0,1200,300),
                       expand=c(0.001,0)) +
  scale_colour_manual(values=mycol,name="Infected") +
    scale_fill_manual(values=mycol,name="Infected")  +
    theme_light()+
    theme(#legend.position="none",
          panel.grid.minor = element_blank()) 
    

landlmmsero<-ggplot(lands, aes(x=day, y=LandDurh ,colour=sero)) + 
    geom_ribbon(data= data.frame(predland), 
            aes(ymin= lower.CL^2, ymax= upper.CL^2, 
                y= NULL,
                fill=sero), 
            alpha=0.4,colour = NA) +
geom_point(alpha=0.6) +
 geom_line(data=predland,aes(x=day,y=emmean^2),lwd=1.5) +
     labs(y="Time in colony (h)",tag="d)",x="") +
    scale_x_continuous(limits=c(0,120),
                     #  labels=daysx,
                       breaks=seq(0,120,20),
                       expand=c(0.001,0.001),
                     labels=c("02/07","22/07","11/08","31/08","20/09","10/10","30/10")) +
    scale_y_continuous(limits=c(0,60),
                       breaks=seq(0,60,15),
                       expand=c(0.001,0)) +
  scale_colour_manual(values=mycol,name="Infected") +
    scale_fill_manual(values=mycol,name="Infected")  +
    theme_light()+
    theme(panel.grid.minor = element_blank()) 
    
mylegendlmmsero<-g_legend(distlmmsero)

tiff(here("outputs","LMM_predictions_sero_2023.tif"),
     width=7000,
     height=4000,
     res=600,
     compression="lzw")
grid.arrange(arrangeGrob(distlmmsero+ theme(legend.position="none"),
                         durlmmsero+ theme(legend.position="none"),
                         pathlmmsero+ theme(legend.position="none"),
                         landlmmsero+ theme(legend.position="none") ,
                                     ncol=2),
               mylegendlmmsero, ncol=2,widths=c(1,0.1)) 

dev.off()

```

Overall, no difference is detected in trip characteristics and nest attendance between deployment 1 an 2. The only significant variable is time

# Statistical tests with sex

```{r}
#| echo: false
#| include: true
#| warning: false
#| message: false

source(here::here("R","stats_functions.R"))
var<-c('day','sex')

#summary of all trips (26 adults)
trips<-trips6 %>% 
       mutate(day=as.numeric(as.Date(DateEnd) - as.Date("2023-07-02")))

lands<-land %>% 
      mutate(day=as.numeric(as.Date(DateEnd) - as.Date("2023-07-02")))


#daysx<-c("03/07","20/08","30/08","09/09","19/09","29/09","20/10")
newdat<-expand.grid(site=unique(trips$site),
                    day=seq(min(trips$day),
                            max(trips$day),1),
                #    sero=unique(trips$sero),
                    sex=unique(trips$sex))

newdatland<-expand.grid(site=unique(lands$site),
                    day=seq(min(lands$day),
                            max(lands$day),1),
             #       sero=unique(lands$sero),
                    sex=unique(lands$sex))

#stats for maximal distance
ggplot(trips, aes(x=day, y=Distmaxkm, 
                  colour=id)) +
    geom_point()+
    geom_smooth(method="lm",alpha=0.2)+
    ggtitle("Maximal distance")


glm.dist<-lme(sqrt(Distmaxkm)~ day + sex,
              random= ~1| id,
              data=trips)

summary(glm.dist)
anova(glm.dist)

qqnorm(resid(glm.dist))
qqline(resid(glm.dist))


#stats for trip duration
ggplot(trips, aes(x=day, y=TripDurh, 
                  colour=id)) +
    geom_point()+
    geom_smooth(method="lm",alpha=0.2)+
    ggtitle("Trip duration")


glm.dur<-lme(sqrt(TripDurh)~ day + sex,
              random= ~1| id,
              data=trips)

summary(glm.dur)
anova(glm.dur)

qqnorm(resid(glm.dur))
qqline(resid(glm.dur))


##stats for total path
ggplot(trips, aes(x=day, y=TotalPathkm, 
                  colour=id)) +
    geom_point()+
    geom_smooth(method="lm",alpha=0.2)+
    ggtitle("Total path")


glm.path<-lme(sqrt(TotalPathkm)~ day + sex,
              random= ~1| id,
              data=trips)

summary(glm.path)
anova(glm.path)

qqnorm(resid(glm.path))
qqline(resid(glm.path))

##stats for time on land
glm.land<-lme(sqrt(LandDurh)~ day + sex,
              random= ~1| id,
              data=lands)

summary(glm.land)
anova(glm.land)

qqnorm(resid(glm.land))
qqline(resid(glm.land))


#predict fitted values and CI

predist<-predict_lme_results(glm.dist,newdat,var)
preddur<-predict_lme_results(glm.dur,newdat,var)
predpath<-predict_lme_results(glm.path,newdat,var)
predland<-predict_lme_results(glm.land,newdatland,var)



distlmmsex<-ggplot(trips, aes(x=day, y=Distmaxkm ,colour=sex)) + 
    geom_ribbon(data= data.frame(predist),
            aes(ymin= lower.CL^2, ymax= upper.CL^2,
                y= NULL,
                fill=sex),
           alpha=0.4,colour = NA) +
geom_point(alpha=0.6) +
 geom_line(data=predist,aes(x=day,y=emmean^2),lwd=1.5) +
     labs(y="Maximal distance to the colony (km)",tag="a)",x="") +
    scale_x_continuous(limits=c(0,120),
                     #  labels=daysx,
                       breaks=seq(0,120,20),
                       expand=c(0.001,0.001)) +
     scale_y_continuous(limits=c(0,400,100),
                        breaks=seq(0,400,100),
                        expand=c(0.001,0.001)) + 
 scale_colour_manual(values=mycol2,name="Sex") +
 scale_fill_manual(values=mycol2,name="Sex")  +
     theme_light()+
    theme(#legend.position="none",
          panel.grid.minor = element_blank()) 

durlmmsex<-ggplot(trips, aes(x=day, y=TripDurh ,colour=sex)) + 
    geom_ribbon(data= data.frame(preddur), 
            aes(ymin= lower.CL^2, ymax= upper.CL^2, 
                y= NULL,
                fill=sex), 
            alpha=0.4,colour = NA) +
geom_point(alpha=0.6) +
 geom_line(data=preddur,aes(x=day,y=emmean^2),lwd=1.5) +
     labs(y="Trip Duration (h)",tag="b)",x="") +
    scale_x_continuous(limits=c(0,120),
                     #  labels=daysx,
                       breaks=seq(0,120,20),
                       expand=c(0.001,0.001)) +
    scale_y_continuous(limits=c(0,200),
                       breaks=seq(0,200,50),
                       expand=c(0.001,0)) +
  scale_colour_manual(values=mycol2,name="Sex") +
    scale_fill_manual(values=mycol2,name="Sex")  +
    theme_light()+
    theme(#legend.position="none",
          panel.grid.minor = element_blank()) 
    

pathlmmsex<-ggplot(trips, aes(x=day, y=TotalPathkm ,colour=sex)) + 
    geom_ribbon(data= data.frame(predpath), 
            aes(ymin= lower.CL^2, ymax= upper.CL^2, 
                y= NULL,
                fill=sex), 
            alpha=0.4,colour = NA) +
geom_point(alpha=0.6) +
 geom_line(data=predpath,aes(x=day,y=emmean^2),lwd=1.5) +
     labs(y="Total distance travelled (km)",tag="c)",x="") +
    scale_x_continuous(limits=c(0,120),
                     #  labels=daysx,
                       breaks=seq(0,120,20),
                       expand=c(0.001,0.001)) +
    scale_y_continuous(limits=c(0,1200),
                       breaks=seq(0,1200,300),
                       expand=c(0.001,0)) +
  scale_colour_manual(values=mycol2,name="Sex") +
    scale_fill_manual(values=mycol2,name="Sex")  +
    theme_light()+
    theme(#legend.position="none",
          panel.grid.minor = element_blank()) 
    

landlmmsex<-ggplot(lands, aes(x=day, y=LandDurh ,colour=sex)) + 
    geom_ribbon(data= data.frame(predland), 
            aes(ymin= lower.CL^2, ymax= upper.CL^2, 
                y= NULL,
                fill=sex), 
            alpha=0.4,colour = NA) +
geom_point(alpha=0.6) +
 geom_line(data=predland,aes(x=day,y=emmean^2),lwd=1.5) +
     labs(y="Time in colony (h)",tag="d)",x="") +
    scale_x_continuous(limits=c(0,120),
                     #  labels=daysx,
                       breaks=seq(0,120,20),
                       expand=c(0.001,0.001)) +
    scale_y_continuous(limits=c(0,60),
                       breaks=seq(0,60,15),
                       expand=c(0.001,0)) +
  scale_colour_manual(values=mycol2,name="Sex") +
    scale_fill_manual(values=mycol2,name="Sex")  +
    theme_light()+
    theme(panel.grid.minor = element_blank()) 
    
mylegendsex<-g_legend(pathlmmsex)

tiff(here("outputs","LMM_predictions_sex_2023.tif"),
     width=7000,
     height=4000,
     res=600,
     compression="lzw")
grid.arrange(arrangeGrob(distlmmsex+ theme(legend.position="none"),
                         durlmmsex+ theme(legend.position="none"),
                         pathlmmsex+ theme(legend.position="none"),
                         landlmmsex+ theme(legend.position="none") ,
                                     ncol=2),
               mylegendsex, ncol=2,widths=c(1,0.1)) 
dev.off()
 
```

Différence entre mâles et femelles pour les trajets en mer mais pas pour l'attendance au nid. Pas d'intéraction avec le statut sérologique.

# Full model with body condition

```{r}

#| echo: false
#| include: true
#| warning: false
#| message: false

source(here::here("R","stats_functions.R"))
var<-c('day','sex','sero')


#daysx<-c("03/07","20/08","30/08","09/09","19/09","29/09","20/10")
newdat<-expand.grid(site=unique(trips.cond$site),
                    day=seq(min(trips.cond$day),
                            max(trips.cond$day),1),
                    sero=unique(trips.cond$sero),
                    sex=unique(trips$sex),
                    body_cond=seq(-2.1,3.3,0.05))

newdatland<-expand.grid(site=unique(lands.cond$site),
                     day=seq(min(lands.cond$day),
                            max(lands.cond$day),1),
                    sero=unique(lands.cond$sero),
                    sex=unique(lands.cond$sex),
                    body_cond=seq(-2.1,3.3,0.05))

glm.dist<-lme(sqrt(Distmaxkm)~ day + sex*sero*body_cond +deploy,
              random= ~1| id,
              data=trips.cond)

summary(glm.dist)
anova(glm.dist)

qqnorm(resid(glm.dist))
qqline(resid(glm.dist))


glm.dur<-lme(sqrt(TripDurh)~ day + sex*sero*body_cond+deploy ,
              random= ~1| id,
              data=trips.cond)

summary(glm.dur)

glm.dur1<-drop1(glm.dur)
anova(glm.dur)

qqnorm(resid(glm.dur))
qqline(resid(glm.dur))


glm.path<-lme(sqrt(TotalPathkm)~ day + sex*sero*body_cond+deploy ,
              random= ~1| id,
              data=trips.cond)

summary(glm.path)
anova(glm.path)

qqnorm(resid(glm.path))
qqline(resid(glm.path))

##stats for time on land
glm<-lme(sqrt(LandDurh)~ day + sex*sero*body_cond+deploy ,
              random= ~1| id,
              data=lands.cond)

summary(glm.land)
anova(glm.land)

qqnorm(resid(glm.land))
qqline(resid(glm.land))


#predict fitted values and CI

predist<-predict_lme_results(glm.dist,newdat,var)
preddur<-predict_lme_results(glm.dur,newdat,var)
predpath<-predict_lme_results(glm.path,newdat,var)

predland<-predict_lme_results(glm.land,newdatland,var)



distlmmsex<-ggplot(trips, aes(x=day, y=Distmaxkm ,colour=sex)) + 
    geom_ribbon(data= data.frame(predist),
            aes(ymin= lower.CL^2, ymax= upper.CL^2,
                y= NULL,
                fill=sex),
           alpha=0.4,colour = NA) +
geom_point(alpha=0.6) +
 geom_line(data=predist,aes(x=day,y=emmean^2),lwd=1.5) +
     labs(y="Maximal distance to the colony (km)",tag="a)",x="") +
    scale_x_continuous(limits=c(0,120),
                     #  labels=daysx,
                       breaks=seq(0,120,20),
                       expand=c(0.001,0.001)) +
     scale_y_continuous(limits=c(0,400,100),
                        breaks=seq(0,400,100),
                        expand=c(0.001,0.001)) + 
 scale_colour_manual(values=mycol2,name="Sex") +
 scale_fill_manual(values=mycol2,name="Sex")  +
     theme_light()+
    theme(#legend.position="none",
          panel.grid.minor = element_blank()) 

durlmmsex<-ggplot(trips, aes(x=day, y=TripDurh ,colour=sex)) + 
    geom_ribbon(data= data.frame(preddur), 
            aes(ymin= lower.CL^2, ymax= upper.CL^2, 
                y= NULL,
                fill=sex), 
            alpha=0.4,colour = NA) +
geom_point(alpha=0.6) +
 geom_line(data=preddur,aes(x=day,y=emmean^2),lwd=1.5) +
     labs(y="Trip Duration (h)",tag="b)",x="") +
    scale_x_continuous(limits=c(0,120),
                     #  labels=daysx,
                       breaks=seq(0,120,20),
                       expand=c(0.001,0.001)) +
    scale_y_continuous(limits=c(0,200),
                       breaks=seq(0,200,50),
                       expand=c(0.001,0)) +
  scale_colour_manual(values=mycol2,name="Sex") +
    scale_fill_manual(values=mycol2,name="Sex")  +
    theme_light()+
    theme(#legend.position="none",
          panel.grid.minor = element_blank()) 
    

pathlmmsex<-ggplot(trips, aes(x=day, y=TotalPathkm ,colour=sex)) + 
    geom_ribbon(data= data.frame(predpath), 
            aes(ymin= lower.CL^2, ymax= upper.CL^2, 
                y= NULL,
                fill=sex), 
            alpha=0.4,colour = NA) +
geom_point(alpha=0.6) +
 geom_line(data=predpath,aes(x=day,y=emmean^2),lwd=1.5) +
     labs(y="Total distance travelled (km)",tag="c)",x="") +
    scale_x_continuous(limits=c(0,120),
                     #  labels=daysx,
                       breaks=seq(0,120,20),
                       expand=c(0.001,0.001)) +
    scale_y_continuous(limits=c(0,1200),
                       breaks=seq(0,1200,300),
                       expand=c(0.001,0)) +
  scale_colour_manual(values=mycol2,name="Sex") +
    scale_fill_manual(values=mycol2,name="Sex")  +
    theme_light()+
    theme(#legend.position="none",
          panel.grid.minor = element_blank()) 
    

landlmmsex<-ggplot(lands, aes(x=day, y=LandDurh ,colour=sex)) + 
    geom_ribbon(data= data.frame(predland), 
            aes(ymin= lower.CL^2, ymax= upper.CL^2, 
                y= NULL,
                fill=sex), 
            alpha=0.4,colour = NA) +
geom_point(alpha=0.6) +
 geom_line(data=predland,aes(x=day,y=emmean^2),lwd=1.5) +
     labs(y="Time in colony (h)",tag="d)",x="") +
    scale_x_continuous(limits=c(0,120),
                     #  labels=daysx,
                       breaks=seq(0,120,20),
                       expand=c(0.001,0.001)) +
    scale_y_continuous(limits=c(0,60),
                       breaks=seq(0,60,15),
                       expand=c(0.001,0)) +
  scale_colour_manual(values=mycol2,name="Sex") +
    scale_fill_manual(values=mycol2,name="Sex")  +
    theme_light()+
    theme(panel.grid.minor = element_blank()) 
    
mylegendsex<-g_legend(pathlmmsex)

tiff(here("outputs","LMM_predictions_sex_2023.tif"),
     width=7000,
     height=4000,
     res=600,
     compression="lzw")
grid.arrange(arrangeGrob(distlmmsex+ theme(legend.position="none"),
                         durlmmsex+ theme(legend.position="none"),
                         pathlmmsex+ theme(legend.position="none"),
                         landlmmsex+ theme(legend.position="none") ,
                                     ncol=2),
               mylegendsex, ncol=2,widths=c(1,0.1)) 
dev.off()
```
