---
title: "2-Trip characteristics"
author: "Aurore Ponchon"
format: html
editor: visual
---

## Trip summary with gaps \<6h

```{r}
library(here)
library(tidyverse)
library(gridExtra)
library(nlme)
library(emmeans)
library(DHARMa)
library(kableExtra)
library(plotrix)

source(here("R","trip_functions.R"))

#trips
load(here::here("data","NewlyCreatedData","clean_6_locs_AD_rouzic23.RData"))
trips6<-trips_summary_ind(loc.clean.interp.rouzic,
                                  colony=colo_coord_rouzic) %>% 
    dplyr::mutate(sero=as.factor(sero))

#time in colony
load(here("data","NewlyCreatedData","rawLocRouzic2023.Rdata"))
land<-raw_rouzic_2023 %>% 
    land_summary_ind(.,colony=colo_coord_rouzic) %>% 
    dplyr::filter(LandDurh > 1 & 
                    LandDurh < 100 & 
                      stage=="adult") %>% 
     dplyr::mutate(sero=as.factor(sero))

#Table summary
sum.ids<-trips6 %>% 
    summarize(Ids=length(unique(id)),
              Trips=length(unique(trip.id)),
              Debut=min(DateEnd),
              Fin=max(DateEnd))
  
kbl(sum.ids,caption="Sample size for trips") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))


sum.id<- trips6 %>% 
    group_by(id) %>% 
    summarize(across(c(Distmaxkm, TripDurh, TotalPathkm), 
                     ~ mean(.x, na.rm = TRUE))) %>% 
    as.data.frame()

sum.trips<- sum.id %>% 
    summarize(across(where(is.numeric), 
                     list(mean=mean,
                          se=plotrix::std.error))) %>% 
    mutate(across(where(is.numeric), round,digits=0))

lands.ids<-land %>% 
    summarize(Ids=length(unique(id)),
              Land=length(unique(onland.id)),
              Debut=min(DateEnd),
              Fin=max(DateEnd))

kbl(lands.ids,caption="Sample size for periods on land") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

land.id<- land %>% 
    group_by(id) %>% 
    summarize(Time= mean(LandDurh, na.rm=T)) %>% 
    as.data.frame()

sum.lands<- land.id %>% 
    summarize(across(where(is.numeric), 
                     list(mean=mean,
                          se=plotrix::std.error))) %>% 
    mutate(across(where(is.numeric), round,digits=0))

kbl(sum.trips,caption="Summary of trip characteristics - Gaps <6h",
    col.names=c("DistmaxMean","DistmaxSE","TripDurMean",
                "TripDurSE","TotPathMean","TotPathSE")) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

kbl(sum.lands,caption="Summary of nest attendance",
    col.names=c("DurationMean","DurationSE")) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

```

## Trip summary with gaps \<3h

```{r}
#trips
load(here::here("data","NewlyCreatedData","clean_3_locs_AD_rouzic23.RData"))
trips3<-trips_summary_ind(loc.clean.interp.rouzic,
                                  colony=colo_coord_rouzic) %>% 
    dplyr::mutate(sero=as.factor(sero))

#Table summary
sum.ids<-trips3 %>% 
    summarize(Ids=length(unique(id)),
              Trips=length(unique(trip.id)),
              Debut=min(DateEnd),
              Fin=max(DateEnd))
  
kbl(sum.ids,caption="Sample size for trips") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))


sum.id<- trips3 %>% 
    group_by(id) %>% 
    summarize(across(c(Distmaxkm, TripDurh, TotalPathkm), 
                     ~ mean(.x, na.rm = TRUE))) %>% 
    as.data.frame()

sum.trips<- sum.id %>% 
    summarize(across(where(is.numeric), 
                     list(mean=mean,
                          se=plotrix::std.error))) %>% 
    mutate(across(where(is.numeric), round,digits=0))


kbl(sum.trips,caption="Summary of trip characteristics - Gaps <3h",
    col.names=c("DistmaxMean","DistmaxSE","TripDurMean",
                "TripDurSE","TotPathMean","TotPathSE")) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))


```

# Comparisons between gaps \<3h and \<6h

```{r}

trips6$Group<-"H6"
trips3$Group<-"H3"
    
trips<-rbind(trips6,trips3)

mycol<-turbo(2,begin=0.25, end=0.7)


 #maximal distance
dist<-ggplot(trips,aes(x=Group,y=Distmaxkm, fill=Group,
                                       colour=Group) ) +
  geom_boxplot(outlier.alpha = 0, alpha=0.3, show.legend = FALSE)+
  geom_jitter(width=0.1, show.legend = FALSE)+
  scale_y_continuous(limits=c(0,400),breaks=seq(0,400,100),expand=c(0.01,0.01))+
 labs(y="Maximal distance to the colony (km)",tag="a)",x="") +
  scale_fill_manual(values=mycol) +
  scale_colour_manual(values=mycol)  +
  theme(text = element_text(size=10),
        axis.text = element_text(size=10),
        legend.text = element_text(size=10),
        axis.title.x=element_blank())

#trip duration
dur<-ggplot(trips,aes(x=Group,y=TripDurh, 
                            fill=Group, colour=Group) )+
  geom_jitter(width=0.1, show.legend = FALSE)+
  geom_boxplot(outlier.alpha = 0, alpha=0.3, show.legend = FALSE)+
  labs(y="Trip duration (hrs)", x="",tag="b)") +
  scale_y_continuous(limits=c(0,200),expand=c(0.01,0.01))+
  scale_fill_manual(values=mycol) +
scale_colour_manual(values=mycol)+
  theme(text = element_text(size=10),
        axis.text = element_text(size=10),
        legend.text = element_text(size=10),
        axis.title.x=element_blank())

#total distance travelled
tot<-ggplot(trips,aes(x=Group,y=TotalPathkm, 
                                      fill=Group, colour=Group) )+
  geom_jitter(width=0.1, show.legend = FALSE)+
  geom_boxplot(outlier.alpha = 0, alpha=0.3, show.legend = FALSE)+
  labs(y="Total distance travelled (km)",x="",tag="c)") +
  scale_y_continuous(limits=c(0,1200),expand=c(0.01,0.01))+
  scale_fill_manual(values=mycol) +
  scale_colour_manual(values=mycol)+
  theme(text = element_text(size=10),
        axis.text = element_text(size=10),
        legend.text = element_text(size=10),
        axis.title.x=element_blank())

#nest attendance
att<-ggplot(land,aes(x=site,y=LandDurh))+
  geom_jitter(width=0.1,show.legend=F)+
  geom_boxplot(outlier.alpha = 0, alpha=0.3, show.legend = FALSE)+
  labs(y="Nest attendance (hrs)",x="",tag="d)") +
#  scale_y_continuous(limits=c(0,120),expand=c(0.01,0.01))+
  scale_fill_manual(values=mycol)  +
  scale_colour_manual(values=mycol)+
  theme(text = element_text(size=10),
        axis.text = element_text(size=10),
        axis.title.x=element_blank())

 # tiff("Figures/Figure 2 - Stats trip characteristics.tif",compression="lzw",
 #      res=600,width=6500,height=3500)
grid.arrange(arrangeGrob(dist,dur,tot,att,
                         ncol=2,nrow=2))



 #maximal distance
distb<-ggplot(trips,aes(x=DateStart,y=Distmaxkm ,group=sero)) +
   geom_point(aes(color=sero,shape=sero),show.legend = F)+

    facet_wrap(Group~., nrow=2) +
 labs(y="Maximal distance to the colony (km)",tag="a)",x="") +
  scale_fill_manual(values=mycol) +
  scale_colour_manual(values=mycol)  

 #max duration
durb<-ggplot(trips,aes(x=DateStart,y=TripDurh ,group=sero)) +
   geom_point(aes(color=sero,shape=sero),show.legend = F)+
    facet_wrap(Group~., nrow=2) +
 labs(y="Trip duration (hrs)",tag="b)",x="") +
  scale_fill_manual(values=mycol) +
  scale_colour_manual(values=mycol)  

 #total distance
totb<-ggplot(trips,aes(x=DateStart,y=TotalPathkm ,
                        group=sero)) +
   geom_point(aes(color=sero,shape=sero),show.legend = F)+
    facet_wrap(Group~., nrow=2) +
 labs(y="Total distance travelled (km)",tag="c)",x="") +
  scale_fill_manual(values=mycol) +
  scale_colour_manual(values=mycol)  

grid.arrange(distb,durb,totb,
             ncol=3)
```

# Comparison between infected and non-infected ind

```{r}
 #maximal distance
trips6$sero<-as.factor(trips6$sero)

dist.sero<-ggplot(trips6,aes(x=sero,y=Distmaxkm, fill=sero,
                                       colour=sero) ) +
  geom_boxplot(outlier.alpha = 0, alpha=0.3, show.legend = FALSE)+
  geom_jitter(width=0.1, show.legend = FALSE) +
  scale_y_continuous(limits=c(0,400),breaks=seq(0,400,100),expand=c(0.01,0.01))+
 labs(y="Maximal distance to the colony (km)",tag="a)",x="") +
  scale_fill_manual(values=mycol) +
  scale_colour_manual(values=mycol)  +
  theme(text = element_text(size=10),
        axis.text = element_text(size=10),
        legend.text = element_text(size=10),
        axis.title.x=element_blank())


#trip duration
dur.sero<-ggplot(trips6,aes(x=sero,y=TripDurh, 
                            fill=sero, colour=sero) )+
  geom_jitter(width=0.1, show.legend = FALSE)+
  geom_boxplot(outlier.alpha = 0, alpha=0.3, show.legend = FALSE)+
  labs(y="Trip duration (hrs)", x="",tag="b)") +
  scale_y_continuous(limits=c(0,200),expand=c(0.01,0.01))+
  scale_fill_manual(values=mycol) +
scale_colour_manual(values=mycol)+
  theme(text = element_text(size=10),
        axis.text = element_text(size=10),
        legend.text = element_text(size=10),
        axis.title.x=element_blank())

#total distance travelled
tot.sero<-ggplot(trips6,aes(x=sero,y=TotalPathkm, 
                                      fill=sero, colour=sero) )+
  geom_jitter(width=0.1, show.legend = FALSE)+
  geom_boxplot(outlier.alpha = 0, alpha=0.3, show.legend = FALSE)+
  labs(y="Total distance travelled (km)",x="",tag="c)") +
  scale_y_continuous(limits=c(0,1200),expand=c(0.01,0.01))+
  scale_fill_manual(values=mycol) +
  scale_colour_manual(values=mycol)+
  theme(text = element_text(size=10),
        axis.text = element_text(size=10),
        legend.text = element_text(size=10),
        axis.title.x=element_blank())

#nest attendance
land<-land %>% 
    dplyr::mutate(Sero=as.factor(sero))

att.sero<-ggplot(land,aes(x=Sero,y=LandDurh,
                     fill=Sero, colour=Sero))+
  geom_jitter(width=0.1,show.legend=F) +
  geom_boxplot(outlier.alpha = 0, alpha=0.3, show.legend = FALSE) +
  labs(y="Nest attendance (hrs)",x="",tag="d)") +
#  scale_y_continuous(limits=c(0,120),expand=c(0.01,0.01))+
  scale_fill_manual(values=mycol)  +
  scale_colour_manual(values=mycol)+
  theme(text = element_text(size=10),
        axis.text = element_text(size=10),
        axis.title.x=element_blank())

 # tiff("Figures/Figure 2 - Stats trip characteristics.tif",compression="lzw",
 #      res=600,width=6500,height=3500)
grid.arrange(arrangeGrob(dist.sero,dur.sero,tot.sero,att.sero,
                         ncol=2,nrow=2))


```
