---
title: "Analysis 2023 - Manuscript"
author: "Aurore Ponchon"
format: html
editor: visual
---

# Trip summary for 2023

The first chunk calculates the trips summary and characteristics and produce summary tables as reported in the manuscript

```{r}
#| echo: false
#| include: true
#| warning: false
#| message: false

library(here)
library(tidyverse)
library(gridExtra)
library(nlme)
library(emmeans)
library(DHARMa)
library(kableExtra)
library(plotrix)
library(readxl)
library(beyonce)
# library(FactoMineR)
# library(corrr)
# library(corrplot)
# library(factoextra)
# library(ggcorrplot)

source(here("R","trip_functions.R"))

mycolsex<-c("#F5D000FF","cyan3") #"deepskyblue3"
mycol2.90<-viridis(n=2,begin=0,end=0.6,option="D")
mycol2.50<-viridis(n=2,begin=0.1,end=0.7,option="D")

#get trips
load(here::here("data","NewlyCreatedData","clean_10_locs_AD_rouzic23.RData"))
trips<-trips_summary_ind(cleanLocs.rouzic23,
                                  colony=colo_coord_rouzic) %>% 
    dplyr::mutate(sero=as.factor(sero),
                   Yeux_colo=as.factor(Yeux_colo), 
                   Yeux_bino=as.factor(Yeux_bino)) %>% 
     mutate(day=as.numeric(as.Date(DateEnd) - as.Date("2023-07-02"))) %>% 
       droplevels()

#time in colony between 2 consecutive trips
load(here("data","NewlyCreatedData","rawLocRouzic2023.Rdata"))
land<-raw_rouzic_2023 %>% 
    land_summary_ind(.,colony=colo_coord_rouzic) %>% 
    dplyr::filter(LandDurh > 1 & 
                    LandDurh < 100 & 
                      stage=="adult") %>% 
     dplyr::mutate(sero=as.factor(sero),
                   Yeux_colo=as.factor(Yeux_colo), 
                   Yeux_bino=as.factor(Yeux_bino)) %>% 
     mutate(day=as.numeric(as.Date(DateEnd) - as.Date("2023-07-02"))) %>% 
      droplevels()

##test to include body condition but not conclusive
# #get the 4 numerical variables which are used as proxies of body condition 
# mat<-trips %>% 
#     dplyr::distinct(id,body_weight,wing, culmen,tarsus, .keep_all = F) %>% 
#     dplyr::select(-id) %>% 
#     scale(.)
# 
# #check correlation between variables
# corr_matrix <- cor(mat)
# corrplot(corr_matrix, method="number")
# 
# #pca to calculate a body condition index
# data.pca <- princomp(mat)
# summary(data.pca)
# fviz_pca_var(data.pca, col.var = "cos2",
#             gradient.cols = c("black", "orange", "green"),
#             repel = TRUE)
# 
# #extract pca results
# cond<-get_pca_ind(data.pca)
# 
# #extract the axis 1 values
# IDS<-trips %>% 
#       dplyr::distinct(id,body_weight,wing, culmen, .keep_all = F) %>% 
#     dplyr::mutate(body_cond=cond$coord[,1])
# 
# #merge body condition index with rest of data
# trips.cond<-trips %>% 
#     left_join(.,IDS,by=c("id","body_weight","culmen","wing"))
# lands.cond<-lands %>% 
#     left_join(.,IDS,by=c("id","body_weight","culmen","wing"))



#Table summary for sample sizes on individuals and trips
sum.ids<-trips %>% 
    group_by(Yeux_bino) %>% 
    summarize(Ids=length(unique(id)),
              Trips=length(unique(trip.id)),
              Debut=min(DateEnd),
              Fin=max(DateEnd)) 
  
kbl(sum.ids,caption="Sample size for trips") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

#summary by individuals
nb.trips.sum<-trips %>% 
    group_by(Yeux_bino,id) %>% 
    summarize(Trips=mean(difftime(max(DateEnd),min(DateStart),units="days"))) %>% 
    summarize(cc=mean(Trips),
              SE=std.error(Trips),
              Min=min(Trips),
              Max=max(Trips))

#count number of trips per individuals
nb.trips<-trips %>% 
     group_by(id, Yeux_bino) %>% 
     count() %>% 
     group_by(Yeux_bino) %>% 
     summarize(Min=min(n),
               Max=max(n))

#calculate mean and sd aby individuals
sum.id<- trips %>% 
    group_by(id,Yeux_bino) %>% 
    summarize(across(c(Distmaxkm, TripDurh, TotalPathkm), 
                     ~ mean(.x, na.rm = TRUE))) %>% 
    as.data.frame()

#calculate mean and se for each iris colour group
sum.trips1<- sum.id %>% 
    group_by(Yeux_bino) %>% 
    summarize(across(where(is.numeric), 
                     list(Mean=mean,
                          SE=plotrix::std.error,
                          Min=min,
                          Max=max))) %>% 
    mutate(across(where(is.numeric), round,digits=1))

#format table
sum.trips<-sum.trips1 %>% 
    group_by(Yeux_bino) %>% 
           dplyr::mutate(MeanDist=paste(Distmaxkm_Mean,Distmaxkm_SE,sep=" ± "),
                         DistRange=paste(Distmaxkm_Min,Distmaxkm_Max,sep=" - "),
                         MeanDur=paste(TripDurh_Mean,TripDurh_SE,sep=" ± "),
                         DurRange=paste(TripDurh_Min,TripDurh_Max,sep=" - "),     
                         MeanPath=paste(TotalPathkm_Mean,TotalPathkm_SE,sep=" ± "),
                      PathRange=paste(TotalPathkm_Min,TotalPathkm_Max,sep=" - ")) %>%
        dplyr::select(-c(2:13))

#summarize data for nest attendance
lands.ids<-land %>% 
    summarize(Ids=length(unique(id)),
              Land=length(unique(onland.id)),
              Debut=min(DateEnd),
              Fin=max(DateEnd))

kbl(lands.ids,caption="Sample size for periods on land") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

#summary by individuals
land.id<- land %>% 
    group_by(Yeux_bino,id) %>% 
    summarize(Time= mean(LandDurh, na.rm=T)) %>% 
    as.data.frame()

#summary by iris colour
sum.lands<- land.id %>% 
    group_by(Yeux_bino) %>% 
    summarize(across(where(is.numeric), 
                     list(mean=mean,
                          se=plotrix::std.error))) %>% 
    mutate(across(where(is.numeric), round,digits=1)) 

#clean table for trips
kbl(sum.trips,caption="Summary of trip characteristics",
    col.names=c("Yeux",
                "Maximal distance (km)",
                "Maximal distance (km)",
                "Trip Duration (h)",
                "Trip Duration (h)",
                "Total path (km)",
                "Total path (km)")) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

#clean table for nest attendance
kbl(sum.lands,caption="Summary of nest attendance",
    col.names=c("Yeux","DurationMean","DurationSE")) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

#calculate mean and se for each group of iris colour
dist.colo<-summarySE(trips, measurevar="Distmaxkm", groupvars=c("Yeux_colo")) %>% 
    mutate(Yeux_colof=fct_relevel(Yeux_colo, "2dark",
                                            "1dark1spotted",
                                            "1dark",
                                            "2spotted",
                                            "1spotted",
                                            "white"))
dur.colo<-summarySE(trips, measurevar="TripDurh", groupvars=c("Yeux_colo"))%>% 
    mutate(Yeux_colof=fct_relevel(Yeux_colo, "2dark",
                                            "1dark1spotted",
                                            "1dark",
                                            "2spotted",
                                            "1spotted",
                                            "white"))
path.colo<-summarySE(trips, measurevar="TotalPathkm", groupvars=c("Yeux_colo"))%>% 
    mutate(Yeux_colof=fct_relevel(Yeux_colo, "2dark",
                                            "1dark1spotted",
                                            "1dark",
                                            "2spotted",
                                            "1spotted",
                                            "white"))
land.colo<-summarySE(land, measurevar="LandDurh", groupvars=c("Yeux_colo"))%>% 
    mutate(Yeux_colof=fct_relevel(Yeux_colo, "2dark",
                                            "1dark1spotted",
                                            "1dark",
                                            "2spotted",
                                            "1spotted",
                                            "white"))

#plot mean and ci of trip characteristics for each iris colour group
col<-beyonce_palette(18)

eyecolordist<-ggplot(dist.colo, aes(x=Yeux_colof, y=Distmaxkm),group=Yeux_colof) +
    geom_point(aes(color=Yeux_colof)) + 
     geom_errorbar(aes(x=Yeux_colof,
                      ymin=Distmaxkm-ci,
                      ymax=Distmaxkm+ci,
                      colour=Yeux_colof),
                  width = 0) +
    scale_y_continuous(limits=c(0,200),
                       expand=c(0.001,0.001))+
    scale_color_manual(values=col, name="Iris colour") +
      labs(y="Maximal distance to the colony (km)",tag="a)",x="") +
    theme_bw()

eyecolordur<-ggplot(dur.colo, aes(x=Yeux_colof, y=TripDurh),group=Yeux_colof) +
     geom_point(aes(color=Yeux_colof)) + 
     geom_errorbar(aes(x=Yeux_colof,
                      ymin=TripDurh-ci,
                      ymax=TripDurh+ci,
                      colour=Yeux_colof),
                  width = 0) +
    scale_y_continuous(limits=c(0,60),
                       expand=c(0.001,0.001)) +
    scale_color_manual(values=col, name="Iris colour") +
      labs(y="Trip duration (h)",tag="b)",x="") +
    theme_bw()

eyecolorpath<-ggplot(path.colo, aes(x=Yeux_colof, y=TotalPathkm),group=Yeux_colof) +
     geom_point(aes(color=Yeux_colof)) + 
     geom_errorbar(aes(x=Yeux_colof,
                      ymin=TotalPathkm-ci,
                      ymax=TotalPathkm+ci,
                      colour=Yeux_colof),
                  width = 0) +
    scale_y_continuous(limits=c(0,600),
                       expand=c(0.001,0.001)) +
    scale_color_manual(values=col, name="Iris colour") +
      labs(y="Total distance travelled (km)",tag="c)",x="") +
    theme_bw()

eyecolorland<-ggplot(land.colo, aes(x=Yeux_colof, y=LandDurh),group=Yeux_colof) +
     geom_point(aes(color=Yeux_colof)) + 
     geom_errorbar(aes(x=Yeux_colof,
                      ymin=LandDurh-ci,
                      ymax=LandDurh+ci,
                      colour=Yeux_colof),
                  width = 0) +
    scale_y_continuous(limits=c(0,30),
                       expand=c(0.001,0.001)) +
    scale_color_manual(values=col, name="Iris colour") +
      labs(y="Nest attendance (h)",tag="a)",x="") +
    theme_bw()

mylegendcoloreye<-g_legend(eyecolordist)

#Figure which is not presented in the study
tiff(here::here("outputs","trips_by_iris_colour.tif"),
     width=7000,
     height=4000,
     res=600,
     compression="lzw")

grid.arrange(arrangeGrob(eyecolordist+ theme(legend.position="none"),
                         eyecolordur+ theme(legend.position="none"),
                         eyecolorpath+ theme(legend.position="none"),
                         eyecolorland+ theme(legend.position="none") ,
                                     ncol=2),
               mylegendcoloreye, ncol=2,widths=c(1,0.1)) 
dev.off()
```

# Statistical analysis 2023

Linear mixed models applied on trip characteristics. Sex, iris colour and their interaction as well as day of the year were tested as fixed effect while bird ID was used as a random effect.

```{r}
#| echo: false
#| include: true
#| warning: false
#| message: false

source(here::here("R","stats_functions.R"))
var<-c('day','sex','Yeux_bino')


#daysx<-c("03/07","20/08","30/08","09/09","19/09","29/09","20/10")
newdat<-expand.grid(site=unique(trips$site),
                    day=seq(min(trips$day),
                            max(trips$day),1),
                    Yeux_bino=unique(trips$Yeux_bino),
                    sex=unique(trips$sex))

newdatland<-expand.grid(site=unique(trips$site),
                    day=seq(min(trips$day),
                            max(trips$day),1),
                    Yeux_bino=unique(trips$Yeux_bino),
                    sex=unique(trips$sex))


glm.dist1<-lme(sqrt(Distmaxkm) ~ day + sex*Yeux_bino ,
              random= ~1| id,
              data=trips)

glm.dist2<-lme(sqrt(Distmaxkm) ~ day + sex+Yeux_bino,
              random= ~1| id,
              data=trips)

glm.dist3<-lme(sqrt(Distmaxkm) ~ day + sex + Yeux_bino,
              random= ~1| id,
              data=trips)

glm.dist4<-lme(sqrt(Distmaxkm) ~ day + sex ,
              random= ~1| id,
              data=trips)

AIC(glm.dist1,glm.dist2,glm.dist3,glm.dist4)


summary(glm.dist1)
anova(glm.dist1)

qqnorm(resid(glm.dist1))
qqline(resid(glm.dist1))


glm.dur1<-lme(sqrt(TripDurh) ~ day + sex*Yeux_bino ,
              random= ~1| id,
              data=trips)

glm.dur2<-lme(sqrt(TripDurh) ~ day + sex+Yeux_bino,
              random= ~1| id,
              data=trips)

glm.dur3<-lme(sqrt(TripDurh) ~ day + sex + Yeux_bino,
              random= ~1| id,
              data=trips)

glm.dur4<-lme(sqrt(TripDurh) ~ day + sex ,
              random= ~1| id,
              data=trips)

AIC(glm.dur1,glm.dur2,glm.dur3,glm.dur4)

summary(glm.dur1)
anova(glm.dur1)

qqnorm(resid(glm.dur1))
qqline(resid(glm.dur1))


##stats for total path
glm.path1<-lme(sqrt(TotalPathkm) ~ day + sex*Yeux_bino ,
              random= ~1| id,
              data=trips)

glm.path2<-lme(sqrt(TotalPathkm) ~ day + sex+Yeux_colo,
              random= ~1| id,
              data=trips)

glm.path3<-lme(sqrt(TotalPathkm) ~ day + sex + Yeux_bino,
              random= ~1| id,
              data=trips)

glm.path4<-lme(sqrt(TotalPathkm) ~ day + sex ,
              random= ~1| id,
              data=trips)

AIC(glm.path1,glm.path2,glm.path3,glm.path4)

summary(glm.path1)
anova(glm.path1)

qqnorm(resid(glm.path1))
qqline(resid(glm.path1))

##stats for time on land
glm.land1<-lme(sqrt(LandDurh)~ day + sex*Yeux_bino,
              random= ~1| id,
              data=land)

glm.land2<-lme(sqrt(LandDurh)~ day + sex+Yeux_bino,
              random= ~1| id,
              data=land)

glm.land3<-lme(sqrt(LandDurh)~ day + sex + Yeux_bino,
              random= ~1| id,
              data=land)

glm.land4<-lme(sqrt(LandDurh)~ day + sex,
              random= ~1| id,
              data=land)

glm.land5<-lme(sqrt(LandDurh)~ day ,
              random= ~1| id,
              data=land)

AIC(glm.land1,glm.land2,glm.land3,glm.land4,glm.land5)


summary(glm.land1)
anova(glm.land1)

qqnorm(resid(glm.land1))
qqline(resid(glm.land1))


#predict fitted values and CI
var<-c('day','sex','Yeux_bino')


predist<-predict_lme_results(glm.dist1,newdat,var)
preddur<-predict_lme_results(glm.dur1,newdat,var)
predpath<-predict_lme_results(glm.path1,newdat,var)
predland<-predict_lme_results(glm.land1,newdatland,var)

mycolinterac<-c("#EF7C12FF","#F5D000FF","#1BB6AFFF","#172869FF")

distlmmsex<-ggplot(trips, aes(x=day, y=Distmaxkm ,colour=sex:Yeux_bino)) + 
    geom_ribbon(data= data.frame(predist),
            aes(ymin= lower.CL^2, ymax= upper.CL^2,
                y= NULL,
                fill=sex:Yeux_bino),
           alpha=0.4,colour = NA) +
geom_point(alpha=0.4) +
 geom_line(data=predist,aes(x=day,y=emmean^2),lwd=1.5) +
     labs(y="Maximal distance to the colony (km)",tag="a)",x="") +
    scale_x_continuous(limits=c(0,110),
                     #  labels=daysx,
                       breaks=seq(0,120,20),
                       expand=c(0.001,0.001),
                     labels=c("02/07","22/07","11/08","31/08","20/09","10/10","30/10")) +
     scale_y_continuous(limits=c(0,400,100),
                        breaks=seq(0,400,100),
                        expand=c(0.001,0.001)) + 
 scale_colour_manual(values=mycolinterac,name="Sex &\niris colour",
                     labels=c('F:0'='F white','F:1'='F black','M:0'='M white','M:1'='M black')) +
 scale_fill_manual(values=mycolinterac,name="Sex &\niris colour",
                     labels=c('F:0'='F white','F:1'='F black','M:0'='M white','M:1'='M black'))  +
     theme_light()+
    theme(#legend.position="none",
          panel.grid.minor = element_blank()) 
print(distlmmsex)

durlmmsex<-ggplot(trips, aes(x=day, y=TripDurh ,colour=sex:Yeux_bino)) + 
    geom_ribbon(data= data.frame(preddur), 
            aes(ymin= lower.CL^2, ymax= upper.CL^2, 
                y= NULL,
                fill=sex:Yeux_bino), 
            alpha=0.4,colour = NA) +
geom_point(alpha=0.4) +
 geom_line(data=preddur,aes(x=day,y=emmean^2),lwd=1.5) +
     labs(y="Trip Duration (h)",tag="b)",x="") +
    scale_x_continuous(limits=c(0,110),
                     #  labels=daysx,
                       breaks=seq(0,120,20),
                       expand=c(0.001,0.001),
                     labels=c("02/07","22/07","11/08","31/08","20/09","10/10","30/10")) +
    scale_y_continuous(limits=c(0,200),
                       breaks=seq(0,200,50),
                       expand=c(0.001,0)) +
 scale_colour_manual(values=mycolinterac,name="Sex &\niris colour",
                     labels=c('F:0'='F white','F:1'='F black','M:0'='M white','M:1'='M black')) +
 scale_fill_manual(values=mycolinterac,name="Sex &\niris colour",
                     labels=c('F:0'='F white','F:1'='F black','M:0'='M white','M:1'='M black'))  +
    theme_light()+
    theme(#legend.position="none",
          panel.grid.minor = element_blank()) 
    

pathlmmsex<-ggplot(trips, aes(x=day, y=TotalPathkm ,colour=sex:Yeux_bino)) + 
    geom_ribbon(data= data.frame(predpath), 
            aes(ymin= lower.CL^2, ymax= upper.CL^2, 
                y= NULL,
                fill=sex:Yeux_bino), 
            alpha=0.4,colour = NA) +
geom_point(alpha=0.4) +
 geom_line(data=predpath,aes(x=day,y=emmean^2),lwd=1.5) +
     labs(y="Total distance travelled (km)",tag="c)",x="") +
    scale_x_continuous(limits=c(0,110),
                     #  labels=daysx,
                       breaks=seq(0,120,20),
                       expand=c(0.001,0.001),
                     labels=c("02/07","22/07","11/08","31/08","20/09","10/10","30/10")) +
    scale_y_continuous(limits=c(0,1200),
                       breaks=seq(0,1200,300),
                       expand=c(0.001,0)) +
 scale_colour_manual(values=mycolinterac,name="Sex &\niris colour",
                     labels=c('F:0'='F white','F:1'='F black','M:0'='M white','M:1'='M black')) +
 scale_fill_manual(values=mycolinterac,name="Sex &\niris colour",
                     labels=c('F:0'='F white','F:1'='F black','M:0'='M white','M:1'='M black'))  +
    theme_light()+
    theme(#legend.position="none",
          panel.grid.minor = element_blank()) 
    

landlmmsex<-ggplot(land, aes(x=day, y=LandDurh ,colour=sex:Yeux_bino)) + 
    geom_ribbon(data= data.frame(predland), 
            aes(ymin= lower.CL^2, ymax= upper.CL^2, 
                y= NULL,
                fill=sex:Yeux_bino), 
            alpha=0.4,colour = NA) +
geom_point(alpha=0.4) +
 geom_line(data=predland,aes(x=day,y=emmean^2),lwd=1.5) +
     labs(y="Nest attendance (h)",tag="d)",x="") +
    scale_x_continuous(limits=c(0,110),
                     #  labels=daysx,
                       breaks=seq(0,120,20),
                       expand=c(0.001,0.001),
                     labels=c("02/07","22/07","11/08","31/08","20/09","10/10","30/10")) +
    scale_y_continuous(limits=c(0,60),
                       breaks=seq(0,60,15),
                       expand=c(0.001,0)) +
 scale_colour_manual(values=mycolinterac,name="Sex &\niris colour",
                     labels=c('F:0'='F white','F:1'='F black','M:0'='M white','M:1'='M black')) +
 scale_fill_manual(values=mycolinterac,name="Sex &\niris colour",
                     labels=c('F:0'='F white','F:1'='F black','M:0'='M white','M:1'='M black'))  +
    theme_light()+
    theme(panel.grid.minor = element_blank()) 
    
mylegendlmmsero<-g_legend(distlmmsex)

tiff(here::here("outputs","Final figures","LMM_predictions_interac_2023.tif"),
     width=7000,
     height=4000,
     res=600,
     compression="lzw")
grid.arrange(arrangeGrob(distlmmsex+ theme(legend.position="none"),
                         durlmmsex+ theme(legend.position="none"),
                         pathlmmsex+ theme(legend.position="none"),
                         landlmmsex+ theme(legend.position="none") ,
                                     ncol=2),
               mylegendlmmsero, ncol=2,widths=c(1,0.1)) 

dev.off()
```

# Spatial analysis 2023

Calculate kernel UD according to sex and iris colour separately, as interaction is not significant.

```{r}
#| echo: false
#| include: true
#| warning: false
#| message: false


library(sf)
library(adehabitatHR)
library(rnaturalearth)
library(terra)
library(raster)
library(ggspatial)
library(RColorBrewer)


source(here::here("R","trip_functions.R"))
source(here::here("R","problem_trips.R"))
source(here::here("R","spatial_function_KUD.R"))

#select locs for 2023
ind.var<-cleanLocs.rouzic23 %>% 
    dplyr::select(id,year,stage,period,sex,
                  age_chick,sero,Yeux_bino,travelNb,trip.id) %>% 
    group_by(id,year) %>% 
    slice(1) %>% 
    dplyr::filter(year==2023) %>% 
    droplevels() %>% 
    ungroup()

#get interpolated data from HMM
interp.23<- read_delim(here::here("data","RawData",
                               "HMM_Rouzic_Morus_bassanus_all_corrected.csv"), 
                  col_names=T,
                  delim = ";",
                  col_types = cols(indID = col_character(), 
                                    datetime = col_datetime(format = 
                                                                "%d/%m/%Y %H:%M"),
                                   longitude=col_number(),
                                   latitude=col_number()), 
                   trim_ws = TRUE) %>% 
    dplyr::rename(long=longitude,lat=latitude) %>% 
    dplyr::mutate(id=as.factor(str_sub(indID,-6,-1))) %>% 
    left_join(.,ind.var,by=c("id","year")) %>% 
    dplyr::filter(year==2023) %>% 
    droplevels() %>% 
    as.data.frame(.) 


#filter and project trips

locs.df23<-interp.23 %>%   #cleanLocs.rouzic23 %>% 
    dplyr::filter(!trip.id %in% migration & 
                      !trip.id %in% last.incomplete) %>% 
    droplevels() %>% 
      sf::st_as_sf(., coords = c("long","lat")) %>%
    sf::st_set_crs(4326) %>%
    st_transform(., projcrs) %>%
   dplyr::mutate(long = sf::st_coordinates(.)[,1],
                 lat = sf::st_coordinates(.)[,2]) %>% 
    as.data.frame(.)


locs.spdf23<-interp.23 %>%  #cleanLocs.rouzic23 %>% 
    dplyr::filter(!trip.id %in% migration & 
                      !trip.id %in% last.incomplete) %>% 
    droplevels() %>%
    sf::st_as_sf(., coords = c("long","lat")) %>%
    sf::st_set_crs(4326) %>%
    st_transform(., projcrs) %>%
    as_Spatial(.)

#rm(cleanLocs.rouzic23)

#load bathymetry layer


#get reduced dataframes 
locs.fem.df23<-get_df_filtered(locs.df23,sex,"F") 
locs.mal.df23<-get_df_filtered(locs.df23,sex,"M") 

locs.pos.df23<-get_df_filtered(locs.df23,Yeux_bino,"1") 
locs.neg.df23<-get_df_filtered(locs.df23,Yeux_bino,"0") 

#get dataframes as ltraj for brb function by sex
locs.ltr.fem23<-as.ltraj(locs.fem.df23[,c("long","lat")], 
             date=locs.fem.df23$datetime, 
             id=locs.fem.df23$id,
             typeII = TRUE)

locs.ltr.mal23<-as.ltraj(locs.mal.df23[,c("long","lat")], 
             date=locs.mal.df23$datetime, 
             id=locs.mal.df23$id,
             typeII = TRUE)

#get dataframes as ltraj for brb function by iris colour
locs.ltr.pos23<-as.ltraj(locs.pos.df23[,c("long","lat")], 
             date=locs.pos.df23$datetime, 
             id=locs.pos.df23$id,
             typeII = TRUE)

locs.ltr.neg23<-as.ltraj(locs.neg.df23[,c("long","lat")], 
             date=locs.neg.df23$datetime, 
             id=locs.neg.df23$id,
             typeII = TRUE)



##kernels with normal ud

# locs.fem.spdf<-subset(locs.spdf , locs.spdf$sex=="F") 
# locs.mal.spdf<-subset(locs.spdf , locs.spdf$sex=="M") 
# KUD.sero.mal<- kernelUD(locs.mal.spdf,
#                     same4all=T,
#                     h=7500,
#                     grid=1000)
# 
# KUD.sero.fem<- kernelUD(locs.fem.spdf,
#                     same4all=T,
#                     h=7500,
#                     grid=1000)
# 
BA.KUD.90<-kerneloverlap(locs.spdf23[,"Yeux_bino"], grid=500,meth="BA",percent=90)
BA.KUD.50<-kerneloverlap(locs.spdf23[,"Yeux_bino"], grid=500,meth="BA",percent=50)
print("Yeux 50 ")
BA.KUD.50
print("Yeux 90")
BA.KUD.90

BA.KUD.90<-kerneloverlap(locs.spdf23[,"sex"], grid=500,meth="BA",percent=90)
BA.KUD.50<-kerneloverlap(locs.spdf23[,"sex"], grid=500,meth="BA",percent=50)
print("Sex 50 ")
      BA.KUD.50
print("Sex 90 ")
      BA.KUD.90


##kernels with brb
TMAX=3600
LMIN=50
HMIN=10000
GRID=500

#get D parameters for BRB function
Tmal23<-BRB.D(locs.ltr.mal23, Tmax = TMAX, Lmin = LMIN)
Tfem23<-BRB.D(locs.ltr.fem23, Tmax = TMAX, Lmin = LMIN)
Tpos23<-BRB.D(locs.ltr.pos23, Tmax = TMAX, Lmin = LMIN)
Tneg23<-BRB.D(locs.ltr.neg23, Tmax = TMAX, Lmin = LMIN)


#BRB function
KUD.mal23<-get_brb_raster(locs.ltr.mal23,Tmal23,
                        TMAX,LMIN,
                        HMIN,GRID)
KUD.fem23<-get_brb_raster(locs.ltr.fem23,Tfem23,
                        TMAX,LMIN,
                        HMIN,GRID)

KUD.pos23<-get_brb_raster(locs.ltr.pos23,Tpos23,
                        TMAX,LMIN,
                        HMIN,GRID)
KUD.neg23<-get_brb_raster(locs.ltr.neg23,Tneg23,
                        TMAX,LMIN,
                        HMIN,GRID)

ver.mal23.90<-st_as_sf(getverticeshr(KUD.mal23, percent=90),crs = projcrs) %>% 
    mutate(Sex="90M")
ver.mal23.50<-st_as_sf(getverticeshr(KUD.mal23, percent=50),crs = projcrs)%>% 
    mutate(Sex="50M")
ver.fem23.90<-st_as_sf(getverticeshr(KUD.fem23, percent=90),crs = projcrs)%>% 
    mutate(Sex="90F")
ver.fem23.50<-st_as_sf(getverticeshr(KUD.fem23, percent=50),crs = projcrs)%>% 
    mutate(Sex="50F")


ver.pos23.90<-st_as_sf(getverticeshr(KUD.pos23, percent=90),crs = projcrs)%>% 
    mutate(Sero="90+")
ver.pos23.50<-st_as_sf(getverticeshr(KUD.pos23, percent=50),crs = projcrs)%>% 
    mutate(Sero="50+")
ver.neg23.90<-st_as_sf(getverticeshr(KUD.neg23, percent=90),crs = projcrs)%>% 
    mutate(Sero="90-")
ver.neg23.50<-st_as_sf(getverticeshr(KUD.neg23, percent=50),crs = projcrs)%>% 
    mutate(Sero="50-")

#BA.KUD.90<-kerneloverlap(locs.spdf[,"sex"], grid=GRID,meth="BA",percent=90)


#extent of study area
    ext.brit<-terra::ext(300000,650000,5270000,5650000)
    
    bathym<-terra::rast(here::here("data",
                                   "RawData",
                                   "shapefiles",
                                   "Bathym_FR_UK.tif")) %>% 
        terra::project(.,projcrs) %>% 
        terra::crop(., ext.brit)
    
    bathym[bathym>0]<-0
    ocean.pal <- rev(colorRampPalette(brewer.pal(9,"Blues"))(9))
    zbreaks <- c(-500,-200,-150,-100,-50,-25,0)
    
    
    #select borders
    target<-c("France","United Kingdom","Spain","Belgium",
              "Netherlands", "Germany","Ireland","Isle of Man",
              "Jersey" ,"Guernsey"  ,"Åland")
    
    ##Plot maps
    #contours of countries
    shp<-ne_countries(scale = 10, returnclass = 'sf') %>%
        dplyr::filter(name %in% target) %>% 
        st_transform(.,projcrs)  %>% 
        st_crop(., ext.brit)
    
    #coastlines
    coast<-ne_coastline(scale = 10, returnclass = "sf") %>% 
        st_transform(.,projcrs)  %>% 
        st_crop(., ext.brit)
    
    #location of colony
    coord_rouzic<-colo_coord_rouzic %>% 
        sf::st_as_sf(., coords = c("long","lat")) %>% 
        sf::st_set_crs(4326) %>% 
        st_transform(.,projcrs)
    

#colors for plot
mycolsero<-c("#ED9C97FF","#E1595CFF","#81C07A", "#004616")
mycolsex<-c("darkorchid1","darkorchid4","#60EFFF","#0093AF") 


#plot
kern.sex23<-ggplot()+
        # tidyterra::geom_spatraster(data = bathym, aes(fill =  Bathym_FR_UK)) +
        geom_sf(data=ver.fem23.90, aes(fill=Sex) , 
                colour=NA,
                alpha=0.2) + 
        geom_sf(data=ver.fem23.50, aes(fill=Sex) , #fill=mycolsex[2],
                colour=NA ,
                alpha=0.7) +
        geom_sf(data=ver.mal23.90, aes(fill=Sex), #fill=mycolsex[3],
                colour=NA,
                alpha=0.5) +
        geom_sf(data=ver.mal23.50, aes(fill=Sex) , #fill=mycolsex[1],
                colour=NA ,
                alpha=0.7) +
        geom_sf(data=shp,fill="grey90") +
        geom_sf(data=coast) +
        geom_sf(data = coord_rouzic,
                shape=23,
                colour="black",
                fill="yellow",
                size=4) +
        coord_sf(xlim=ext.brit[c(1,2)],
                 ylim=ext.brit[c(3,4)],
                 expand=F) +
        labs(tag="b)") +
        scale_fill_manual(values=c("50F"="darkorchid4",
                                   "50M"="#0093AF",
                                   "90F"="darkorchid1",
                                   "90M"="#60EFFF"),
                            labels=c("90F"="90% UD female",
                                     "50F"="50% UD female",
                                     "90M" = "90% UD male",
                                     "50M" = "50% UD male"),
                       #   levels=c("90F","50F","90M","50M"),
                            name="Sex") +
        annotation_scale(location = "br",
                         width_hint = 0.15,
                         pad_x = unit(0.7, "cm"),
                         bar_cols = "black") +
        annotation_north_arrow(location = "br",
                               which_north = "true",
                               style = ggspatial::north_arrow_nautical(),
                               pad_y = unit(0.8, "cm")) +
        theme_bw() +
      theme(legend.position="bottom")+
      guides(fill=guide_legend(ncol=2))
   
  
print(kern.sex23)



#plot
  kern.sero23<-ggplot()+
   # tidyterra::geom_spatraster(data = bathym, aes(fill =  Bathym_FR_UK)) +
     geom_sf(data=ver.pos23.90, aes(fill=Sero) , 
                colour=NA,
                alpha=0.4) + 
        geom_sf(data=ver.pos23.50, aes(fill=Sero) , #fill=mycolsex[2],
                colour=NA ,
                alpha=0.7) +
        geom_sf(data=ver.neg23.90, aes(fill=Sero), #fill=mycolsex[3],
                colour=NA,
                alpha=0.4) +
        geom_sf(data=ver.neg23.50, aes(fill=Sero) , #fill=mycolsex[1],
                colour=NA ,
                alpha=0.5) +
   geom_sf(data=shp,fill="grey90") +
  geom_sf(data=coast) +
 geom_sf(data = coord_rouzic,
         shape=23,                   
         colour="black",
         fill="yellow",
         size=4) +
 coord_sf(xlim=ext.brit[c(1,2)],
          ylim=ext.brit[c(3,4)],
          expand=F) +  
  labs(tag="a)") +
        scale_fill_manual(values=c("50+"="#E1595CFF",
                                   "50-"="#004616",
                                   "90+"="#ED9C97FF",
                                   "90-"="#81C07A"),
                            labels=c("90+"="90% UD black",
                                     "50+"="50% UD black",
                                     "90-" = "90% UD white",
                                     "50-" = "50% UD white"),
                       #   levels=c("90F","50F","90M","50M"),
                            name="Iris\ncolour") +
    annotation_scale(location = "br",
                     width_hint = 0.15,
                     pad_x = unit(0.7, "cm"),
                     bar_cols = "black") +
    annotation_north_arrow(location = "br",
                           which_north = "true",
                           style = ggspatial::north_arrow_nautical(),
                           pad_y = unit(0.8, "cm")) +
    theme_bw()+
      theme(legend.position="bottom")+
      guides(fill=guide_legend(ncol=2))
  
  print(kern.sero23)

   tiff(here::here("outputs","Final figures",
                   "adults_distribution_brb_2023_final_legend.tif"),
     width=4000,
     height=2500,
     res=500,
     compression="lzw")

grid.arrange(kern.sero23,kern.sex23,ncol=2,nrow=1)

dev.off()


```

# 
